### 2017-05-04 paper.tex
First commit

\documentclass[10pt]{sensys-proc}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}
\graphicspath{{./figures/} {./plots/}}

\usepackage{balance}
\usepackage{comment}
\usepackage{xspace}
\usepackage[export]{adjustbox}
\usepackage{wrapfig}
\newcommand{\chain}{Chain\xspace}
%\numberofauthors{1}

%\author{
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
%% e-mail address with \email.
%\alignauthor Alice Security \\
%        \affaddr{Department of Computer Science}\\
%        \affaddr{University of Southern California}\\
%       \email{alice@example.edu}
%\alignauthor Bob Privacy \\
%    \affaddr{Networked Embedded Systems Group}\\
%    \affaddr{Swedish Institute of Computer Science}\\
%    \email{bob@example.se}
%}

\title{Chains on Threads on Tasks on Channels....}

\begin{document}

\maketitle

%Only including an abstract since it looks like the submission site wants one... 
\begin{abstract}
Killer abstract. 
\end{abstract}

\section{Introduction/Background}
  \label{sec:intro}
Intro stuff about intermittence that Brandon already knows but we should record for the
heck of it. Here's a sample citation for totally unrelated things~\cite{RC,Grace}!

%Sample of adding a picture across the top of a split page... 
\begin{figure*}
\centering
\begin{minipage}[b]{0.49\textwidth}
  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
\caption{Capybara power system overview}\label{label-a}
\end{minipage}\hfill
\begin{minipage}[b]{0.49\textwidth}
  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
\caption{Effect of different power modes}\label{label-b}
\end{minipage}
\end{figure*}

More text after the picture to show how that looks...

\section{Implementation}
We only had to break \chain a little. We promise!

\section{Results}
No really, there will be things here!

\section{Conclusion}
Multithreading in intermittence {\em is} a good idea!

\balance
\bibliographystyle{abbrv}
\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
\end{document}

### 2017-05-09 paper.tex
Trailing whitespace fixes, #authors = 3
Changed font size to 11, added gitignore
Added stuff we whiteboarded to the tex file

@@
-\documentclass[10pt]{sensys-proc}
+\documentclass[11pt]{sensys-proc}
 
@@
 \newcommand{\chain}{Chain\xspace}
-%\numberofauthors{1}
+\numberofauthors{3}
 
@@
 
-\title{Chains on Threads on Tasks on Channels....}
+\title{TAPIR: Threaded Application Programming on Intermittant Resources}
 
@@
   \label{sec:intro}
-Intro stuff about intermittence that Brandon already knows but we should record for the
-heck of it. Here's a sample citation for totally unrelated things~\cite{RC,Grace}!
+  Chain, DINO\\
+  Multithreading\\
+  Intermittance Review\\
+  RMW\\
+  Example Citation\cite{RC,Grace}!
 
@@
 
+\section{Related Work}
+Operating system work\\
+TU delft guys...\\
+Energy Align\\
+
+\section{Chain Discussion}
+Arrays in channels and why that's bad\\
+MAX\_NUM\_DIRTY\_FIELDS\\
+Explioting the fact that there isn't a compiler\\
+
+\section{What we did} % TODO change title
+Scheduler\\
+Interrupts\\
+Mutexes\\
+User programming model for threads\\
+Interrupts\\
+Round Robin (alternatives, on reboot)\\
+SIMD vs MIMD\\
+SIMD w/ channels (you don't get your own registers...etc -
+    just channels, since intermittant)\\
+
 \section{Implementation}
 We only had to break \chain a little. We promise!
+Scheduler Data structures and idempotence (what we broke and why we broke it)\\
+Mutexes (Simple, meant for pseudo channels)\\
+Interrupts (mucked with IE flags)\\
 
@@
 No really, there will be things here!
+Applications/measurements\\
+Blinker (x2)\\
+RSA+cukoo\\
+No verification\\
+Memory overhead (read off compiliation)\\
+Performance (\# reboots)\\
+
+\section{Future Work}
+Duplicating channels (a little renaming or sharing arrays with mutexes)\\
+Optimizations to reduce overhead\\
 

### 2017-05-10 paper.tex
Added lots of TODOs, Interrupts under "what we did" mostly done

--- 2017-05-04 paper.tex
@@
 \usepackage{comment}
+\usepackage{listings}
 \usepackage{xspace}
@@
 
-\title{Chains on Threads on Tasks on Channels....}
+\title{TAPIR: Threaded Application Programming on Intermittent Resources}
 
@@
 
-\section{Introduction/Background}
+\section{Introduction/Background} -Amanda
   \label{sec:intro}
-Intro stuff about intermittence that Brandon already knows but we should record for the
-heck of it. Here's a sample citation for totally unrelated things~\cite{RC,Grace}!
+  Chain, DINO\\
+  Multithreading\\
+  Intermittence Review\\
+  RMW\\
+  Example Citation\cite{Chain, Mementos, Aware, Dino}!
 
@@
 
+\section{Related Work} -Emily
+Operating system work\\
+TU delft guys...\\
+Energy Align\\
+
+\section{Chain Discussion} -NEIL
+
+Arrays in channels and why that's bad\\
+MAX\_NUM\_DIRTY\_FIELDS\\
+Exploiting the fact that there isn't a compiler\\
+
+\section{What we did} % TODO change title
+Scheduler\\
+Mutexes - Emily\\
+User programming model for threads - Neil\\
+Round Robin (alternatives, on reboot)\\
+SIMD vs MIMD\\
+SIMD w/ channels (you don't get your own registers...etc -
+    just channels, since intermittent)\\
+
+\subsection{Interrupts}
+\subsubsection{Motivation}
+Ideally, multithreading would extend to a programming model that can utilize
+interrupts in a producer/consumer fashion (i.e. a thread is waiting on data
+from an external source, the external source fires an interrupt, the thread
+samples the external source and stores the received data in a buffer, and
+the thread returns to its original task). However, the inherent
+non-deterministic
+nature of interrupts from, say, an ADC, creates additional complexity for the
+programmer when reasoning about shared memory, especially with respect to
+idempotence.\\
+    Specific to the MSP430 architecture (but common in other systems), there is
+also the issue of stack integrity. On the MSP430, when an interrupt fires,
+the program counter and status register are pushed onto the stack for
+the handler to be able to return back to the state of the machine when the
+interrupt fired. In our case, however, where power failure is the common
+case, an application could experience failure while inside an interrupt
+handler. This raises two issues. First, upon reboot from a power failure,
+the stack will be cleared (since the stack is stored in volatile memory) and
+execution will return to the handler, if we assume that the interrupt handler
+is a \chain task. Upon leaving the interrupt, via the MSP430 RETI instruction),
+the program's behavior will be undefined. Second, if the programmer does not
+use a \chain task for the interrupt handler, it is possible that the interrupt
+could fire when the system has a low amount of energy remaining, thus
+rendering the system unable to finish executing the interrupt handler,
+leading to a host of potential problems.
+
+\subsubsection{Interface}
+We provide the functions and macros in figure 3 as an interface to
+the programmer. The standard workflow is as follows: the programmer
+designates the task that should be run on an interrupt firing
+with \texttt{INTERRUPT\_TASK(val, func)} where \texttt{val} is the task
+index and \texttt{func} is the function that will be called upon an
+interrupt firing - \texttt{val} and \texttt{func} follow the same
+interface as \chain tasks. Since we assume that the handler is not
+reentrant, the programmer replaces calls to \texttt{\_\_enable\_interrupt()}
+with our function \texttt{enable\_interrupts()}.
+The programmer's code remains unmodified, with the exception of a call to
+\texttt{int\_setup\_complete()} when the \chain program has finished whatever
+register initialization is necessary to setup the interrupts for the
+application.\\
+
+\begin{figure*}
+\begin{minipage}[b]{0.49\textwidth}
+\begin{lstlisting}
+    /** Enable interrupts if int_setup_complete() has been called
+      * and execution is not inside an interrupt handler */
+    void enable_interrupts();
+
+    /** To be called when setup for the interrupt handler is complete,
+      * also enables interrupts */
+    void int_setup_complete();
+
+    /** Return from the interrupt handler */
+    void return_from_interrupt();
+
+    /** Designate a task to run when an interrupt fires */
+    INTERRUPT_TASK(val, func)
+\end{lstlisting}
+\caption{Interrupt Handling Interface}\label{label-a}
+\end{minipage}\hfill
+\end{figure*}
+
+\subsubsection{Design}
+Working within the constraints of \chain and our multi-programming extensions,
+we work to provide a layer of abstraction to the programmer through the standard
+layer of tasks and channels. Notably, this includes interrupt handlers being
+tasks, rather than standard functions. The alternative has been explored
+by keeping specific capacitors for interrupt handlers\cite{Aware}, but this
+approach fails if interrupt handlers take longer than the capacitor has
+energy for. This bug handlers not completing due to power failure is
+unintuitive for programmers coming from consistently powered devices and
+makes certain devices unusable under intermittent power.\\
+
+With interrupt handlers as tasks, if a power failure occurs during a
+handler, execution will resume at the handler. This isn't ideal, the
+standard paradigm that interrupt handlers are generally short and fast
+so that interrupts won't be missed (for instance, a timer interrupt
+being missed because processing the interrupt took longer than the
+timer period), but we argue that it is more unreasonable to say
+that interrupt handlers that take too long will not complete than
+to violate the standard paradigm, since the former arbitrarily restricts
+the application domain.\\
+
+For the sake of simplicity, we assume that interrupt handlers are not
+reentrant. Since reentrant handlers can be used in a non-reentrant fashion
+without consequences, this assumption does not restrict the application domain.
+Moreover, the benefit to reentrant handlers, the application will not miss an
+interrupt firing because processing the previous one took too long, is
+secondary on intermittent architectures - there is likely a higher chance that
+a power failure will occur than an interrupt handler causing an interrupt to be
+missed, unless interrupts are firing on an extremely fine granularity.\\
+
+When an interrupt handler has completed, the programmer calls
+\texttt{return\_from\_interrupt()}, which returns execution to the start
+of the task that was running when the interrupt occurred. This contradicts
+expected behavior - on continuously powered systems, execution generally
+returns to the instruction after the instruction that the system received an
+interrupt while processing. The continuous system model is intuitive, but
+breaks under \chain's intermittence. Task idempotence in \chain only works
+when execution returns to the top of a
+task upon power failure. Program state (registers and stack memory) are stored
+in volatile memory, if an interrupt handler returned to the middle of a task,
+it would be impossible to reason about program correctness, since none of the
+program state had been stored. The programmer could store program state on the
+stack at the start of interrupt handlers, but this would be lost on power
+failure. She could also store this program state in non-volatile memory -
+trivial for the register set, but difficult and potentially extremely costly to
+store the stack, as execution could be several functions deep. Even determining
+how much stack to save could be prohibitively costly. Another alternative would
+be to store program state in non-volatile, but this has been shown to be
+prohibitively costly in most applications \cite{Aware}. \\
+
+Interrupts - NEIL\\
+
 \section{Implementation}
 We only had to break \chain a little. We promise!
+Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
+Mutexes (Simple, meant for pseudo channels) - Emily\\
+Interrupts (mucked with IE flags) - Neil\\
+% Interrupt prologue
+% Interrupt setup
 
-\section{Results}
+\section{Results} - Emily
 No really, there will be things here!
+Applications/measurements\\
+Blinker (x2)\\
+RSA+cukoo\\
+No verification\\
+Memory overhead (read off compilation)\\
+Performance (\# reboots)\\
+
+\section{Future Work} -Amanda
+Duplicating channels (a little renaming or sharing arrays with mutexes)\\
+Optimizations to reduce overhead\\
 
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
+\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography
 \end{document}

### 2017-05-10 paper.tex
outline for sections

--- 2017-05-09 paper.tex
@@
   \label{sec:intro}
-  Chain, DINO\\
-  Multithreading\\
-  Intermittance Review\\
+  Chain, DINO <-- THIS ME\\
+  Multithreading <-- THIS ME\\
+  Intermittance Review <-- THIS ME\\
   RMW\\
@@
 Optimizations to reduce overhead\\
+\\
+Automatically renaming channels to reuse task code\\
+-challenging to define sharing between threads\\
+-challenging to constrain memory for channels (not that much of it, also compilation)\\
+-challenge to uniquely identify multiple channel instances of the same task\\
+\\
+Optimizations to reduce overhead\\
+-convert free indices array to bitmap\\
+-compile-time re-assignation of re-used channels\\
 

### 2017-05-10 paper.tex

--- 2017-05-10 paper.tex
@@
   Multithreading\\
-  Intermittence Review\\
+  Intermittance Review\\
   RMW\\

### 2017-05-10 paper.tex
future work

--- 2017-05-10 paper.tex
@@
 
-\section{Future Work} -Amanda
-Duplicating channels (a little renaming or sharing arrays with mutexes)\\
-Optimizations to reduce overhead\\
+\section{Future Work}
+Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks. 
+When forking a new task, a channel may be duplicated, which linearly increases the amount of memory neceesary for channels with the number of currently running programs.
+It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how 
+many are to be running simulateously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
+In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
+Currently channels are allocated and named at compile time by combining the names of the input and output channel tasks. 
+However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
+This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.\\
+
+Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
+To do this the concept of interrmittent tasks mutexes for shared resources and data structures would be introduced.
+The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
+In order to allow all tasks access to a data structure without breaking all chain semantics, a modified version of a module channel would be used in order 
+for multiple tasks to potentially modify one channel. 
+However, even with that change it is challenging to define sharing between threads in an intermittent context and to maintain chain semantics while implementing 
+shared data structures while still maintaining correctness. 
+Because channel sizes are constrained at compile time currently, it is also necessary to know at compile time all data structures, shared or private, 
+that are going to be used during execution of a chain program or programs. 
+It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory 
+is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.\\
+
+Future work also includes optimizations to reduce overhead, such as converting the free indices array to a free indices bitmap in order to reduce the amount of 
+baseline memory that the scheduler task uses. 
+Re-assignation and re-use of channels at runtime would increaes overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
+If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.\\
 

### 2017-05-10 paper.tex
fixed whitespace in future work

@@
 However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
-This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.\\
+This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.
 
@@
 It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory 
-is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.\\
+is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.
 
@@
 Re-assignation and re-use of channels at runtime would increaes overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
-If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.\\
+If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.
 

### 2017-05-10 paper.tex
Commented out kapybera pictures, added Chain discussion, fixed indentation issues

--- 2017-05-10 paper.tex
@@
   Multithreading\\
-  Intermittance Review\\
+  Intermittence Review\\
   RMW\\
@@
 %Sample of adding a picture across the top of a split page...
-\begin{figure*}
-\centering
-\begin{minipage}[b]{0.49\textwidth}
-  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
-\caption{Capybara power system overview}\label{label-a}
-\end{minipage}\hfill
-\begin{minipage}[b]{0.49\textwidth}
-  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
-\caption{Effect of different power modes}\label{label-b}
-\end{minipage}
-\end{figure*}
+
+%\begin{figure*}
+%\centering
+%\begin{minipage}[b]{0.49\textwidth}
+%  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
+%\caption{Capybara power system overview}\label{label-a}
+%\end{minipage}\hfill
+%\begin{minipage}[b]{0.49\textwidth}
+%  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
+%\caption{Effect of different power modes}\label{label-b}
+%\end{minipage}
+%\end{figure*}
 
@@
 
-\section{Chain Discussion} -NEIL
+\section{Chain Discussion}
+\chain, while providing a reasonable interface for programmers to work within,
+has several inefficiencies that give poor performance on certain applications.
+First, there is no notion of data sharing between tasks - all communication is
+strictly done through channels. This is useful, as the programmer never needs
+to reason about idempotence with respect to their data, but all data that needs
+to be transferred between tasks needs to be housed in a channel. Since, under
+the original constraints of \chain, channels cannot be written or read from by
+tasks that do not own the channel - only tasks that own the channel can access
+the data within. This means that large data structures, arrays for instance,
+that need to be shared between several channels must be unnecessarily
+duplicated in non-volatile memory - one copy of the data structure for each
+channel that it needs to be accessed in. Granted, there is no enforcement of
+channel ownership by the compiler, but the programmer risks idempotence
+violations by straying outside the constraints of chain. We exploit this to
+create an equivalent to the multicast-in and multicast-out channels described
+in \chain, but using them requires much more reasoning on the part of the programmer.
 
-Arrays in channels and why that's bad\\
-MAX\_NUM\_DIRTY\_FIELDS\\
-Exploiting the fact that there isn't a compiler\\
+Additionally, since dynamic memory allocation is arguably too expensive to run
+on intermittently powered systems, memory is allocated at compile time.
+Unfortunately, since there isn't a compiler specifically for \chain, arrays
+are statically sized based on \texttt{\#define} parameters. This can be
+create tricky memory corruption errors for the programmer when out-of-bounds
+array accesses occur because, for instance, a self-channel has more dirty
+fields than \texttt{MAX\_NUM\_DIRTY\_FIELDS}.
+
+Future work for \chain needs to include some sort of memory allocation based on
+the set of channels allocated in the program, but this is work for a compiler
+and beyond the scope of this work. For our purposes of threaded applications,
+however, \chain is a sufficient launching pad, despite its misgivings since it
+does not require specialized hardware or energy subsystems to work effectively,
+and since it solves the issue of idempotent data accesses in a way that is
+abstracted from the programmer.
 
@@
 makes certain devices unusable under intermittent power.\\
-
 With interrupt handlers as tasks, if a power failure occurs during a
@@
 to violate the standard paradigm, since the former arbitrarily restricts
-the application domain.\\
+the application domain.
 
@@
 a power failure will occur than an interrupt handler causing an interrupt to be
-missed, unless interrupts are firing on an extremely fine granularity.\\
+missed, unless interrupts are firing on an extremely fine granularity.
 
@@
 be to store program state in non-volatile, but this has been shown to be
-prohibitively costly in most applications \cite{Aware}. \\
-
-Interrupts - NEIL\\
+prohibitively costly in most applications \cite{Aware}.
 
@@
 Optimizations to reduce overhead\\
-\\
-Automatically renaming channels to reuse task code\\
--challenging to define sharing between threads\\
--challenging to constrain memory for channels (not that much of it, also compilation)\\
--challenge to uniquely identify multiple channel instances of the same task\\
-\\
-Optimizations to reduce overhead\\
--convert free indices array to bitmap\\
--compile-time re-assignation of re-used channels\\
 

### 2017-05-10 paper.tex

@@
 
-\section{Future Work} -Amanda
-Duplicating channels (a little renaming or sharing arrays with mutexes)\\
-Optimizations to reduce overhead\\
+\section{Future Work}
+Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks. 
+When forking a new task, a channel may be duplicated, which linearly increases the amount of memory neceesary for channels with the number of currently running programs.
+It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how 
+many are to be running simulateously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
+In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
+Currently channels are allocated and named at compile time by combining the names of the input and output channel tasks. 
+However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
+This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.
+
+Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
+To do this the concept of interrmittent tasks mutexes for shared resources and data structures would be introduced.
+The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
+In order to allow all tasks access to a data structure without breaking all chain semantics, a modified version of a module channel would be used in order 
+for multiple tasks to potentially modify one channel. 
+However, even with that change it is challenging to define sharing between threads in an intermittent context and to maintain chain semantics while implementing 
+shared data structures while still maintaining correctness. 
+Because channel sizes are constrained at compile time currently, it is also necessary to know at compile time all data structures, shared or private, 
+that are going to be used during execution of a chain program or programs. 
+It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory 
+is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.
+
+Future work also includes optimizations to reduce overhead, such as converting the free indices array to a free indices bitmap in order to reduce the amount of 
+baseline memory that the scheduler task uses. 
+Re-assignation and re-use of channels at runtime would increaes overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
+If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.
 

### 2017-05-10 paper.tex
small change

--- 2017-05-10 paper.tex
@@
 \usepackage{comment}
-\usepackage{listings}
 \usepackage{xspace}
@@
 
-\section{Introduction/Background} -Amanda
+\section{Introduction/Background}
   \label{sec:intro}
@@
   Multithreading\\
-  Intermittence Review\\
+  Intermittance Review\\
   RMW\\
-  Example Citation\cite{Chain, Mementos, Aware, Dino}!
+  Example Citation\cite{RC,Grace}!
 
 %Sample of adding a picture across the top of a split page...
-
-%\begin{figure*}
-%\centering
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
-%\caption{Capybara power system overview}\label{label-a}
-%\end{minipage}\hfill
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
-%\caption{Effect of different power modes}\label{label-b}
-%\end{minipage}
-%\end{figure*}
+\begin{figure*}
+\centering
+\begin{minipage}[b]{0.49\textwidth}
+  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
+\caption{Capybara power system overview}\label{label-a}
+\end{minipage}\hfill
+\begin{minipage}[b]{0.49\textwidth}
+  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
+\caption{Effect of different power modes}\label{label-b}
+\end{minipage}
+\end{figure*}
 
@@
 
-\section{Related Work} -Emily
+\section{Related Work}
 Operating system work\\
@@
 Energy Align\\
+TinyOS stuff in wireless sensor networks\\
 
 \section{Chain Discussion}
-\chain, while providing a reasonable interface for programmers to work within,
-has several inefficiencies that give poor performance on certain applications.
-First, there is no notion of data sharing between tasks - all communication is
-strictly done through channels. This is useful, as the programmer never needs
-to reason about idempotence with respect to their data, but all data that needs
-to be transferred between tasks needs to be housed in a channel. Since, under
-the original constraints of \chain, channels cannot be written or read from by
-tasks that do not own the channel - only tasks that own the channel can access
-the data within. This means that large data structures, arrays for instance,
-that need to be shared between several channels must be unnecessarily
-duplicated in non-volatile memory - one copy of the data structure for each
-channel that it needs to be accessed in. Granted, there is no enforcement of
-channel ownership by the compiler, but the programmer risks idempotence
-violations by straying outside the constraints of chain. We exploit this to
-create an equivalent to the multicast-in and multicast-out channels described
-in \chain, but using them requires much more reasoning on the part of the programmer.
-
-Additionally, since dynamic memory allocation is arguably too expensive to run
-on intermittently powered systems, memory is allocated at compile time.
-Unfortunately, since there isn't a compiler specifically for \chain, arrays
-are statically sized based on \texttt{\#define} parameters. This can be
-create tricky memory corruption errors for the programmer when out-of-bounds
-array accesses occur because, for instance, a self-channel has more dirty
-fields than \texttt{MAX\_NUM\_DIRTY\_FIELDS}.
-
-Future work for \chain needs to include some sort of memory allocation based on
-the set of channels allocated in the program, but this is work for a compiler
-and beyond the scope of this work. For our purposes of threaded applications,
-however, \chain is a sufficient launching pad, despite its misgivings since it
-does not require specialized hardware or energy subsystems to work effectively,
-and since it solves the issue of idempotent data accesses in a way that is
-abstracted from the programmer.
+Arrays in channels and why that's bad\\
+MAX\_NUM\_DIRTY\_FIELDS\\
+Explioting the fact that there isn't a compiler\\
 
@@
 Scheduler\\
-Mutexes - Emily\\
-User programming model for threads - Neil\\
+Interrupts\\
+Mutexes\\
+User programming model for threads\\
+Interrupts\\
 Round Robin (alternatives, on reboot)\\
@@
 SIMD w/ channels (you don't get your own registers...etc -
-    just channels, since intermittent)\\
-
-\subsection{Interrupts}
-\subsubsection{Motivation}
-Ideally, multithreading would extend to a programming model that can utilize
-interrupts in a producer/consumer fashion (i.e. a thread is waiting on data
-from an external source, the external source fires an interrupt, the thread
-samples the external source and stores the received data in a buffer, and
-the thread returns to its original task). However, the inherent
-non-deterministic
-nature of interrupts from, say, an ADC, creates additional complexity for the
-programmer when reasoning about shared memory, especially with respect to
-idempotence.\\
-    Specific to the MSP430 architecture (but common in other systems), there is
-also the issue of stack integrity. On the MSP430, when an interrupt fires,
-the program counter and status register are pushed onto the stack for
-the handler to be able to return back to the state of the machine when the
-interrupt fired. In our case, however, where power failure is the common
-case, an application could experience failure while inside an interrupt
-handler. This raises two issues. First, upon reboot from a power failure,
-the stack will be cleared (since the stack is stored in volatile memory) and
-execution will return to the handler, if we assume that the interrupt handler
-is a \chain task. Upon leaving the interrupt, via the MSP430 RETI instruction),
-the program's behavior will be undefined. Second, if the programmer does not
-use a \chain task for the interrupt handler, it is possible that the interrupt
-could fire when the system has a low amount of energy remaining, thus
-rendering the system unable to finish executing the interrupt handler,
-leading to a host of potential problems.
-
-\subsubsection{Interface}
-We provide the functions and macros in figure 3 as an interface to
-the programmer. The standard workflow is as follows: the programmer
-designates the task that should be run on an interrupt firing
-with \texttt{INTERRUPT\_TASK(val, func)} where \texttt{val} is the task
-index and \texttt{func} is the function that will be called upon an
-interrupt firing - \texttt{val} and \texttt{func} follow the same
-interface as \chain tasks. Since we assume that the handler is not
-reentrant, the programmer replaces calls to \texttt{\_\_enable\_interrupt()}
-with our function \texttt{enable\_interrupts()}.
-The programmer's code remains unmodified, with the exception of a call to
-\texttt{int\_setup\_complete()} when the \chain program has finished whatever
-register initialization is necessary to setup the interrupts for the
-application.\\
-
-\begin{figure*}
-\begin{minipage}[b]{0.49\textwidth}
-\begin{lstlisting}
-    /** Enable interrupts if int_setup_complete() has been called
-      * and execution is not inside an interrupt handler */
-    void enable_interrupts();
-
-    /** To be called when setup for the interrupt handler is complete,
-      * also enables interrupts */
-    void int_setup_complete();
-
-    /** Return from the interrupt handler */
-    void return_from_interrupt();
-
-    /** Designate a task to run when an interrupt fires */
-    INTERRUPT_TASK(val, func)
-\end{lstlisting}
-\caption{Interrupt Handling Interface}\label{label-a}
-\end{minipage}\hfill
-\end{figure*}
-
-\subsubsection{Design}
-Working within the constraints of \chain and our multi-programming extensions,
-we work to provide a layer of abstraction to the programmer through the standard
-layer of tasks and channels. Notably, this includes interrupt handlers being
-tasks, rather than standard functions. The alternative has been explored
-by keeping specific capacitors for interrupt handlers\cite{Aware}, but this
-approach fails if interrupt handlers take longer than the capacitor has
-energy for. This bug handlers not completing due to power failure is
-unintuitive for programmers coming from consistently powered devices and
-makes certain devices unusable under intermittent power.\\
-With interrupt handlers as tasks, if a power failure occurs during a
-handler, execution will resume at the handler. This isn't ideal, the
-standard paradigm that interrupt handlers are generally short and fast
-so that interrupts won't be missed (for instance, a timer interrupt
-being missed because processing the interrupt took longer than the
-timer period), but we argue that it is more unreasonable to say
-that interrupt handlers that take too long will not complete than
-to violate the standard paradigm, since the former arbitrarily restricts
-the application domain.
-
-For the sake of simplicity, we assume that interrupt handlers are not
-reentrant. Since reentrant handlers can be used in a non-reentrant fashion
-without consequences, this assumption does not restrict the application domain.
-Moreover, the benefit to reentrant handlers, the application will not miss an
-interrupt firing because processing the previous one took too long, is
-secondary on intermittent architectures - there is likely a higher chance that
-a power failure will occur than an interrupt handler causing an interrupt to be
-missed, unless interrupts are firing on an extremely fine granularity.
-
-When an interrupt handler has completed, the programmer calls
-\texttt{return\_from\_interrupt()}, which returns execution to the start
-of the task that was running when the interrupt occurred. This contradicts
-expected behavior - on continuously powered systems, execution generally
-returns to the instruction after the instruction that the system received an
-interrupt while processing. The continuous system model is intuitive, but
-breaks under \chain's intermittence. Task idempotence in \chain only works
-when execution returns to the top of a
-task upon power failure. Program state (registers and stack memory) are stored
-in volatile memory, if an interrupt handler returned to the middle of a task,
-it would be impossible to reason about program correctness, since none of the
-program state had been stored. The programmer could store program state on the
-stack at the start of interrupt handlers, but this would be lost on power
-failure. She could also store this program state in non-volatile memory -
-trivial for the register set, but difficult and potentially extremely costly to
-store the stack, as execution could be several functions deep. Even determining
-how much stack to save could be prohibitively costly. Another alternative would
-be to store program state in non-volatile, but this has been shown to be
-prohibitively costly in most applications \cite{Aware}.
+    just channels, since intermittant)\\
 
@@
 We only had to break \chain a little. We promise!
-Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
-Mutexes (Simple, meant for pseudo channels) - Emily\\
-Interrupts (mucked with IE flags) - Neil\\
-% Interrupt prologue
-% Interrupt setup
+Scheduler Data structures and idempotence (what we broke and why we broke it)\\
+Mutexes (Simple, meant for pseudo channels)\\
+Interrupts (mucked with IE flags)\\
 
-\section{Results} - Emily
+\section{Results}
 No really, there will be things here!
@@
 No verification\\
-Memory overhead (read off compilation)\\
+Memory overhead (read off compiliation)\\
 Performance (\# reboots)\\
 
-\section{Future Work} -Amanda
+\section{Future Work}
 Duplicating channels (a little renaming or sharing arrays with mutexes)\\
@@
 \bibliographystyle{abbrv}
-\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography
+\bibliography{sigproc}  % sigproc.bib is the name of the Bibliography in this case
 \end{document}

### 2017-05-10 paper.tex

--- 2017-05-10 paper.tex
@@
 %Sample of adding a picture across the top of a split page...
-
-%\begin{figure*}
-%\centering
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
-%\caption{Capybara power system overview}\label{label-a}
-%\end{minipage}\hfill
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
-%\caption{Effect of different power modes}\label{label-b}
-%\end{minipage}
-%\end{figure*}
+\begin{figure*}
+\centering
+\begin{minipage}[b]{0.49\textwidth}
+  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
+\caption{Capybara power system overview}\label{label-a}
+\end{minipage}\hfill
+\begin{minipage}[b]{0.49\textwidth}
+  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
+\caption{Effect of different power modes}\label{label-b}
+\end{minipage}
+\end{figure*}
 
@@
 Energy Align\\
+TinyOS stuff in wireless sensor networks\\
 
-\section{Chain Discussion}
-\chain, while providing a reasonable interface for programmers to work within,
-has several inefficiencies that give poor performance on certain applications.
-First, there is no notion of data sharing between tasks - all communication is
-strictly done through channels. This is useful, as the programmer never needs
-to reason about idempotence with respect to their data, but all data that needs
-to be transferred between tasks needs to be housed in a channel. Since, under
-the original constraints of \chain, channels cannot be written or read from by
-tasks that do not own the channel - only tasks that own the channel can access
-the data within. This means that large data structures, arrays for instance,
-that need to be shared between several channels must be unnecessarily
-duplicated in non-volatile memory - one copy of the data structure for each
-channel that it needs to be accessed in. Granted, there is no enforcement of
-channel ownership by the compiler, but the programmer risks idempotence
-violations by straying outside the constraints of chain. We exploit this to
-create an equivalent to the multicast-in and multicast-out channels described
-in \chain, but using them requires much more reasoning on the part of the programmer.
+\section{Chain Discussion} -NEIL
 
-Additionally, since dynamic memory allocation is arguably too expensive to run
-on intermittently powered systems, memory is allocated at compile time.
-Unfortunately, since there isn't a compiler specifically for \chain, arrays
-are statically sized based on \texttt{\#define} parameters. This can be
-create tricky memory corruption errors for the programmer when out-of-bounds
-array accesses occur because, for instance, a self-channel has more dirty
-fields than \texttt{MAX\_NUM\_DIRTY\_FIELDS}.
-
-Future work for \chain needs to include some sort of memory allocation based on
-the set of channels allocated in the program, but this is work for a compiler
-and beyond the scope of this work. For our purposes of threaded applications,
-however, \chain is a sufficient launching pad, despite its misgivings since it
-does not require specialized hardware or energy subsystems to work effectively,
-and since it solves the issue of idempotent data accesses in a way that is
-abstracted from the programmer.
+Arrays in channels and why that's bad\\
+MAX\_NUM\_DIRTY\_FIELDS\\
+Exploiting the fact that there isn't a compiler\\
 
@@
 makes certain devices unusable under intermittent power.\\
+
 With interrupt handlers as tasks, if a power failure occurs during a
@@
 to violate the standard paradigm, since the former arbitrarily restricts
-the application domain.
+the application domain.\\
 
@@
 a power failure will occur than an interrupt handler causing an interrupt to be
-missed, unless interrupts are firing on an extremely fine granularity.
+missed, unless interrupts are firing on an extremely fine granularity.\\
 
@@
 be to store program state in non-volatile, but this has been shown to be
-prohibitively costly in most applications \cite{Aware}.
+prohibitively costly in most applications \cite{Aware}. \\
+
+Interrupts - NEIL\\
 

### 2017-05-11 paper.tex
introduction first draft

@@
 
-\section{Introduction/Background} -Amanda
+\section{Introduction/Background}
   \label{sec:intro}
-  Chain, DINO\\
-  Multithreading\\
-  Intermittance Review\\
-  RMW\\
+%A/N: I don't think that the paragraphs in this section are in a great order, but I am having a hard time making it flow better.
+%Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
+%Also also: I'm trying to get [energy-harvesting computing devices] and [intermittent execution model] below to be in italics,
+%but when I use \textit it makes the ENTIRE PAPER in italics, i used the curly braces and everything helpppppp that has never happened to me before
+
+The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
+New technologies in both of these areas has led to new [energy-harvesting computing devices] that operates only with power it has 
+collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.
+
+Commonly, an energy-harvesting devices charges and collects power in a storage capacitor until the capacitor reaches a sufficient level of charge.
+Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
+This leads to the [intermittent execution model] where regular shut downs and reboots to recharge are the common case rather than the rare case in
+a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave
+irregular behavior on an intermittent device and a new programming model is needed.
+
+Our multithreaded model is build on two existing programming models, DINO and Chain.
+An initial intermittent programming model DINO breaks intermittent programs into tasks, which creates automatic checkpointing locations 
+during exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks both volatile and nonvolatile memory.
+The DINO execution model executes all tasks transactionally, leading to atomic task semantics and does not permit consistency violations in 
+nonvolatile state \cite{Dino}. Chain, a more recent iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses task boundaries
+called channels to store most used data in nonvolatile memory \cite{Chain}. At compile time, Chain allocates the necessary channels between tasks
+by parsing channel read and write statements in order to match up aligning tasks. All inputs and outputs to each task, as well as necessary data
+structures, are stored in a channel for a respective task. A channel may only be modified by one task, and only one task may run at a time,
+which mitigates time of check, time of use bugs (TOCTOU) in intermittence. 
+The one exception is a self-channel for recursive code, which is double buffered.
+Keeping all modifiable data in nonvolatile memory that is not open to easy rewrite makes these tasks idempotent across reboots.
+Because of continuous reboots, many operations necessarily will be executed more than once,
+and correctness is compromised if those operations are not guaranteed to have the same outcome by the programming model after each reboot.
+
+A main issue in intermittent computing along with idempotence is the Read-Modify-Write (RMW) problem. 
+Because volatile memory is cleared on each reboot, it is possible that in between reading and modifying or modifying and writing the data had changed,
+leading to incorrect answers and an inconsistent state. In the Chain programming model, this problem is solved by creating a new task to house the
+problematic operations. Because tasks are blocked by channels that can only be written by one task and are saved to nonvolatile memory, RMW is avoided.
+However, in multiprogramming the problem manifests in another way; if the scheduler switches tasks on a reboot, the currently running thread could have
+been in the middle of its critical section, making it difficult for other threads to make progress if context switches always occur after a reboot.
+Because of this we have decided on a round-robin scheduler with context switches at the end of a task, to guarantee forward progress while knowing
+which tasks have run to completion.
+
+In the current Chain programming model, writing code that completes multiple objectives correctly is now difficult. 
+Current chain programs are small and only have one objective. Additionally, there is no support for advanced control flow.
+Multiple control threads may exist, but in the current model there is no support for them and programmability is difficult and often incorrect.
+Supporting multiple control threads in Chain also necessitates a linear increase in a limited memory resource, because all large shared data structures
+must be copied into each necessary channel, and also makes updating the structures a time consuming, high-overhead task.
+In order to increase functionality of chain and to mitigate these shortcomings, we add familiar threading semantics to Chain in order to express
+multi-threaded control flow operation. In order to do this we create a lightweight scheduler task that becomes part of the inter-task checkpointing
+operation and provide a method of sharing large data structures between tasks with a locking feature. This method is low-memory and low-execution
+overhead, which can then maintain the Chain programming semantics at the user level still successfully complete work on an intermittent device.
+
 Example Citation\cite{Chain, Mementos, Aware, Dino}!
@@
 Energy Align\\
-TinyOS stuff in wireless sensor networks\\
 

### 2017-05-11 paper.tex
italics

--- 2017-05-11 paper.tex
@@
 Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
-This leads to the [intermittent execution model] where regular shut downs and reboots to recharge are the common case rather than the rare case in
+This leads to the \textit{intermittent execution model} where regular shut downs and reboots to recharge are the common case rather than the rare case in
 a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave

### 2017-05-11 paper.tex
Fixed issue with italics, removed trailing whitespace from intro and future notes

--- 2017-05-11 paper.tex
@@
 %Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
-%Also also: I'm trying to get [energy-harvesting computing devices] and [intermittent execution model] below to be in italics,
-%but when I use \textit it makes the ENTIRE PAPER in italics, i used the curly braces and everything helpppppp that has never happened to me before
 
 The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
-New technologies in both of these areas has led to new [energy-harvesting computing devices] that operates only with power it has 
+New technologies in both of these areas has led to new
+\textit{energy-harvesting computing devices} that operates only with power it has
 collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.

### 2017-05-11 paper.tex
updated references and related work

--- 2017-05-11 paper.tex
@@
 
-\section{Introduction/Background}
+\section{Introduction/Background} -Amanda
   \label{sec:intro}
-%A/N: I don't think that the paragraphs in this section are in a great order, but I am having a hard time making it flow better.
-%Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
-%Also also: I'm trying to get [energy-harvesting computing devices] and [intermittent execution model] below to be in italics,
-%but when I use \textit it makes the ENTIRE PAPER in italics, i used the curly braces and everything helpppppp that has never happened to me before
-
-The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
-New technologies in both of these areas has led to new [energy-harvesting computing devices] that operates only with power it has 
-collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.
-
-Commonly, an energy-harvesting devices charges and collects power in a storage capacitor until the capacitor reaches a sufficient level of charge.
-Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
-This leads to the \textit{intermittent execution model} where regular shut downs and reboots to recharge are the common case rather than the rare case in
-a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave
-irregular behavior on an intermittent device and a new programming model is needed.
-
-Our multithreaded model is build on two existing programming models, DINO and Chain.
-An initial intermittent programming model DINO breaks intermittent programs into tasks, which creates automatic checkpointing locations 
-during exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks both volatile and nonvolatile memory.
-The DINO execution model executes all tasks transactionally, leading to atomic task semantics and does not permit consistency violations in 
-nonvolatile state \cite{Dino}. Chain, a more recent iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses task boundaries
-called channels to store most used data in nonvolatile memory \cite{Chain}. At compile time, Chain allocates the necessary channels between tasks
-by parsing channel read and write statements in order to match up aligning tasks. All inputs and outputs to each task, as well as necessary data
-structures, are stored in a channel for a respective task. A channel may only be modified by one task, and only one task may run at a time,
-which mitigates time of check, time of use bugs (TOCTOU) in intermittence. 
-The one exception is a self-channel for recursive code, which is double buffered.
-Keeping all modifiable data in nonvolatile memory that is not open to easy rewrite makes these tasks idempotent across reboots.
-Because of continuous reboots, many operations necessarily will be executed more than once,
-and correctness is compromised if those operations are not guaranteed to have the same outcome by the programming model after each reboot.
-
-A main issue in intermittent computing along with idempotence is the Read-Modify-Write (RMW) problem. 
-Because volatile memory is cleared on each reboot, it is possible that in between reading and modifying or modifying and writing the data had changed,
-leading to incorrect answers and an inconsistent state. In the Chain programming model, this problem is solved by creating a new task to house the
-problematic operations. Because tasks are blocked by channels that can only be written by one task and are saved to nonvolatile memory, RMW is avoided.
-However, in multiprogramming the problem manifests in another way; if the scheduler switches tasks on a reboot, the currently running thread could have
-been in the middle of its critical section, making it difficult for other threads to make progress if context switches always occur after a reboot.
-Because of this we have decided on a round-robin scheduler with context switches at the end of a task, to guarantee forward progress while knowing
-which tasks have run to completion.
-
-In the current Chain programming model, writing code that completes multiple objectives correctly is now difficult. 
-Current chain programs are small and only have one objective. Additionally, there is no support for advanced control flow.
-Multiple control threads may exist, but in the current model there is no support for them and programmability is difficult and often incorrect.
-Supporting multiple control threads in Chain also necessitates a linear increase in a limited memory resource, because all large shared data structures
-must be copied into each necessary channel, and also makes updating the structures a time consuming, high-overhead task.
-In order to increase functionality of chain and to mitigate these shortcomings, we add familiar threading semantics to Chain in order to express
-multi-threaded control flow operation. In order to do this we create a lightweight scheduler task that becomes part of the inter-task checkpointing
-operation and provide a method of sharing large data structures between tasks with a locking feature. This method is low-memory and low-execution
-overhead, which can then maintain the Chain programming semantics at the user level still successfully complete work on an intermittent device.
-
+  Chain, DINO\\
+  Multithreading\\
+  Intermittance Review\\
+  RMW\\
   Example Citation\cite{Chain, Mementos, Aware, Dino}!
@@
 \section{Related Work} -Emily
+In addition to the division of a program into tasks, as described above, prior work
+explores several different strategies for preserving state across reboots. The approaches
+range from fully independent of the programmer~\cite{ratchet, dewdrop} at one extreme, to programmer
+defined task boundaries~\cite{dino} at the other.  Compiler driven approaches such as the
+method used in ~\cite{Mementos} allow the programmer to specify high level behavior of the
+checkpointing system. A key similarity among the prior work is it
+does not lend itself well to multi-programming let
+alone multi-threading. In fact, many of these works state the assumption that the device
+hardware cannot support a conventional OS, and thus focus on single threaded programs. 
+Despite the availability of (relatively) inexpensive nonvolatile memory technology such as
+FRAM~\cite{quickrecall}, the the energy and storage overhead of checkpointing multiple
+thread states would limit the maximum per-thread checkpoint size.
+
+Furthermore, if a multiprogrammed application were implemented using previous
+checkpointing strategies, reasoning about thread interleavings becomes  difficult to
+reason about because of unpredictable power failures. Any conventional timeout based
+thread switching approach will be challenging in an intermittent system, but systems that
+allows dynamic checkpointing~\cite{hibernus}, or dynamic task formation~\cite{Dino} become
+even more difficult to program. Particularly in systems like ~\cite{Dino} where the
+programmer is responsible for placing task boundaries, there is a risk that programmers
+will either overuse task boundaries in an effort to preserve any given thread's state, or
+underprovision task boundaries such that progress is never made because of context
+switching that dynamically expands the number of instructions between task boundaries.
+
+While prior work has stopped short of suggesting an operating system for intermittently
+powered devices, a number of different works touch on the need for these devices to
+perform multiple tasks in the same workload. Buettner, Greenstein and Wetherall motivate
+their runtime, Dewdrop, based on difference in energy requirements for the tasks that a
+CRFID needs to perform~\cite{dewdrop}. The energy aware memory mapping in ~\cite{Aware}
+discusses the challenge of handling nondeterministic interrupts received by intermittent
+sensing platforms when a peripheral must be serviced immediately and control flow deviates
+from single threaded execution. From a hardware standpoint, the federated energy scheme in
+~\cite{ufop} provides a platform for removing the energy dependence between an MCU and its
+peripherals. Federated energy also simplifies a programmer's reasoning about reading from
+a single peripheral. However, federating energy does not help a programmer describe how to
+handle {\em multiple} peripherals. Consider a listing in
+~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
+of the hardware behavior a programmer would have to employ to, say, read one sensor if it
+is available but read from and proceed to process data from a different sensor if it is not.
+
 Operating system work\\
+Want to do multiple things- i.e. federated energy \\
+Dewdrop- energy aware task scheduling\\
 TU delft guys...\\
 Energy Align\\
+TinyOS stuff in wireless sensor networks\\
 
@@
 SIMD vs MIMD\\
-SIMD w/ channels (you don't get your own registers...etc -
-    just channels, since intermittent)\\
-
+SIMD w/ channels (you don't get your registers...etc -
+    just channels, since intermittent\\
 \subsection{Interrupts}

### 2017-05-11 paper.tex
Added more related works

@@
 \DeclareGraphicsExtensions{.pdf,.jpg,.png}
-\graphicspath{{./figures/} {./plots/}}
+\graphicspath{{./figs/} {./plots/}}
 
@@
 %\author{
-%
+
 % The command \alignauthor (no curly braces needed) should
@@
 % affiliation/address with \affaddr, and tag the
-%% e-mail address with \email.
-%\alignauthor Alice Security \\
-%        \affaddr{Department of Computer Science}\\
-%        \affaddr{University of Southern California}\\
-%       \email{alice@example.edu}
-%\alignauthor Bob Privacy \\
-%    \affaddr{Networked Embedded Systems Group}\\
-%    \affaddr{Swedish Institute of Computer Science}\\
-%    \email{bob@example.se}
+% e-mail address with \email.
+
+%\alignauthor Amanda Marano \\
+%        \affaddr{Department of Electrical and Computer Engineering}\\
+%        \affaddr{Carnegie Mellon University}\\
+%       \email{amarano@andrew.cmu.edu}
+%\alignauthor Emily Ruppel \\
+%        \affaddr{Department of Electrical and Computer Engineering}\\
+%        \affaddr{Carnegie Mellon University}\\
+%       \email{eruppel@andrew.cmu.edu}
+%\alignauthor Neil Ryan \\
+%        \affaddr{Department of Electrical and Computer Engineering}\\
+%        \affaddr{Carnegie Mellon University}\\
+%       \email{nryan@andrew.cmu.edu}
 %}
@@
 
-Operating system work\\
-Want to do multiple things- i.e. federated energy \\
-Dewdrop- energy aware task scheduling\\
-TU delft guys...\\
-Energy Align\\
-TinyOS stuff in wireless sensor networks\\
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.99\columnwidth]{figs/listing}
+  \caption{{\bf Example program for sensing and sending data}}
+  \label{fig:copy}
+\end{figure}
+
+The wireless sensor networks community has explored multiple approaches for handling
+multitasking on resource limited devices. The designs tend to provide a combination of
+timer and event driven context switching as well as a simple programming interface.
+TinyOS uses a simple FIFO scheduler to handle event based context switches~\cite{tiny}.
+Nano-RK implements priority based preemptive multitasking where task priorities are
+statically defined~\cite{nano}. Contiki uses proto-theads, stackless program threads, to
+offer a model between nano-RK and tinyOS. The Contiki implements "cooperative
+multithreading" such that events can trigger a context switch, but multithreading may also
+be timer based~\cite{contiki}. The goal of multitasking interfaces in wireless sensor
+networks is to provide a tradeoff between network lifetime and quality of service. This
+differs from the intermittent programming model where the challenge is not system
+lifetime, but system functionality based on energy availability.
 
@@
 Multithreading in intermittence {\em is} a good idea!
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figs/tapir}
+\end{figure}
 

### 2017-05-11 paper.tex
Fixed missing citations

@@
 
-\section{Related Work} -Emily
+\section{Related Work} 
 In addition to the division of a program into tasks, as described above, prior work
@@
 range from fully independent of the programmer~\cite{ratchet, dewdrop} at one extreme, to programmer
-defined task boundaries~\cite{dino} at the other.  Compiler driven approaches such as the
+defined task boundaries~\cite{Dino} at the other.  Compiler driven approaches such as the
 method used in ~\cite{Mementos} allow the programmer to specify high level behavior of the

### 2017-05-11 paper.tex
Added section on long-range control flow, added section on scheduler and programming model. Spelling fixes in future work

--- 2017-05-11 paper.tex
@@
 \section{Related Work} -Emily
-In addition to the division of a program into tasks, as described above, prior work
-explores several different strategies for preserving state across reboots. The approaches
-range from fully independent of the programmer~\cite{ratchet, dewdrop} at one extreme, to programmer
-defined task boundaries~\cite{dino} at the other.  Compiler driven approaches such as the
-method used in ~\cite{Mementos} allow the programmer to specify high level behavior of the
-checkpointing system. A key similarity among the prior work is it
-does not lend itself well to multi-programming let
-alone multi-threading. In fact, many of these works state the assumption that the device
-hardware cannot support a conventional OS, and thus focus on single threaded programs. 
-Despite the availability of (relatively) inexpensive nonvolatile memory technology such as
-FRAM~\cite{quickrecall}, the the energy and storage overhead of checkpointing multiple
-thread states would limit the maximum per-thread checkpoint size.
-
-Furthermore, if a multiprogrammed application were implemented using previous
-checkpointing strategies, reasoning about thread interleavings becomes  difficult to
-reason about because of unpredictable power failures. Any conventional timeout based
-thread switching approach will be challenging in an intermittent system, but systems that
-allows dynamic checkpointing~\cite{hibernus}, or dynamic task formation~\cite{Dino} become
-even more difficult to program. Particularly in systems like ~\cite{Dino} where the
-programmer is responsible for placing task boundaries, there is a risk that programmers
-will either overuse task boundaries in an effort to preserve any given thread's state, or
-underprovision task boundaries such that progress is never made because of context
-switching that dynamically expands the number of instructions between task boundaries.
-
-While prior work has stopped short of suggesting an operating system for intermittently
-powered devices, a number of different works touch on the need for these devices to
-perform multiple tasks in the same workload. Buettner, Greenstein and Wetherall motivate
-their runtime, Dewdrop, based on difference in energy requirements for the tasks that a
-CRFID needs to perform~\cite{dewdrop}. The energy aware memory mapping in ~\cite{Aware}
-discusses the challenge of handling nondeterministic interrupts received by intermittent
-sensing platforms when a peripheral must be serviced immediately and control flow deviates
-from single threaded execution. From a hardware standpoint, the federated energy scheme in
-~\cite{ufop} provides a platform for removing the energy dependence between an MCU and its
-peripherals. Federated energy also simplifies a programmer's reasoning about reading from
-a single peripheral. However, federating energy does not help a programmer describe how to
-handle {\em multiple} peripherals. Consider a listing in
-~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
-of the hardware behavior a programmer would have to employ to, say, read one sensor if it
-is available but read from and proceed to process data from a different sensor if it is not.
-
 Operating system work\\
-Want to do multiple things- i.e. federated energy \\
-Dewdrop- energy aware task scheduling\\
 TU delft guys...\\
 Energy Align\\
-TinyOS stuff in wireless sensor networks\\
 
@@
 
+\chain also struggles with long-range control flow. For each transition to a
+new task, the transition is an explicit one-to-one mapping. \chain provides no
+semantics, for instance, to transfer control from task A to task B, then also
+transfer control to task C after task B completes, without the programmer
+placing \texttt{TRANSITION\_TO(task\_C)} inside task B. This could be done by
+using channels in a similar manner to function pointers, while staying within
+the semantics of \chain (as we did), but reasoning about code becomes more
+challenging for the programmer.
+
 Future work for \chain needs to include some sort of memory allocation based on
@@
 \section{What we did} % TODO change title
-Scheduler\\
 Mutexes - Emily\\
-User programming model for threads - Neil\\
 Round Robin (alternatives, on reboot)\\
 SIMD vs MIMD\\
-SIMD w/ channels (you don't get your registers...etc -
-    just channels, since intermittent\\
+SIMD w/ channels (you don't get your own registers...etc -
+    just channels, since intermittent)\\
+
+\subsection{Scheduler}
+For a multi-threaded programming model to exist, one must create a structure to
+switch between multiple threads. To keep with \chain's model, we chose to make
+the scheduler its own task. The alternative is possible, but since our
+implementation of the scheduler required multicast channels, it makes more
+sense to have the multicast channels have one endpoint well defined, instead of
+some sort of all-to-all channel (which is just a global location in
+non-volatile memory). The multicast channels channels are not strictly
+necessary, but, all in all, we believe that keeping the scheduler as an
+explicit task is good for abstraction. Additionally, if a task has few
+self-channels and the data structures passed in those self-channels are
+relatively small, the overhead of transitioning to that task is relatively
+small, as few double-buffers for self-channels need to be flipped (which is
+arguably the largest overhead).
+
+\subsection{Programming Model}
+Threaded applications that use POSIX or some other standard library interface
+tend to use function calls to interact with the library - for instance, the
+user calls a function \texttt{thread\_create} to spawn off a thread.
+Unfortunately, this is a difficult programming model to keep under
+intermittence, since function calls like \texttt{thread\_create} tend to
+involve non-idempotent actions, like appending to a list or incrementing a
+variable's value. This leaves two clear options - making the functions to
+interact with our threading library \chain tasks, or adding infrastructure to
+to our threading library to make the non-idempotent actions that need to occur
+act idempotently. We opted for the latter - we feel that it is more valuable to
+maintain programmability by provide the user with an interface that they
+expect, instead of trying to extend a \chain semantic like
+\texttt{TRANSITION\_TO} to also include thread creation.
+
 \subsection{Interrupts}
@@
 programmer when reasoning about shared memory, especially with respect to
-idempotence.\\
+idempotence.
+
     Specific to the MSP430 architecture (but common in other systems), there is
@@
 \begin{figure*}
-\begin{minipage}[b]{0.49\textwidth}
+\begin{minipage}[b]{1.0\textwidth}
 \begin{lstlisting}
@@
 unintuitive for programmers coming from consistently powered devices and
-makes certain devices unusable under intermittent power.\\
+makes certain devices unusable under intermittent power.
+
 With interrupt handlers as tasks, if a power failure occurs during a
@@
 \section{Implementation}
-We only had to break \chain a little. We promise!
 Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
@@
 Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks.
-When forking a new task, a channel may be duplicated, which linearly increases the amount of memory neceesary for channels with the number of currently running programs.
+When forking a new task, a channel may be duplicated, which linearly increases the amount of memory necessary for channels with the number of currently running programs.
 It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how
-many are to be running simulateously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
+many are to be running simultaneously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
 In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
@@
 Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
-To do this the concept of interrmittent tasks mutexes for shared resources and data structures would be introduced.
+To do this the concept of intermittent tasks mutexes for shared resources and data structures would be introduced.
 The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
@@
 baseline memory that the scheduler task uses.
-Re-assignation and re-use of channels at runtime would increaes overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
+Re-assignation and re-use of channels at runtime would increase overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
 If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.

### 2017-05-11 paper.tex

@@
 
-\section{Related Work} -Emily
-Operating system work\\
-TU delft guys...\\
-Energy Align\\
+\section{Related Work} 
+In addition to the division of a program into tasks, as described above, prior work
+explores several different strategies for preserving state across reboots. The approaches
+range from fully independent of the programmer~\cite{ratchet, dewdrop} at one extreme, to programmer
+defined task boundaries~\cite{Dino} at the other.  Compiler driven approaches such as the
+method used in ~\cite{Mementos} allow the programmer to specify high level behavior of the
+checkpointing system. A key similarity among the prior work is it
+does not lend itself well to multi-programming let
+alone multi-threading. In fact, many of these works state the assumption that the device
+hardware cannot support a conventional OS, and thus focus on single threaded programs. 
+Despite the availability of (relatively) inexpensive nonvolatile memory technology such as
+FRAM~\cite{quickrecall}, the the energy and storage overhead of checkpointing multiple
+thread states would limit the maximum per-thread checkpoint size.
+
+Furthermore, if a multiprogrammed application were implemented using previous
+checkpointing strategies, reasoning about thread interleavings becomes  difficult to
+reason about because of unpredictable power failures. Any conventional timeout based
+thread switching approach will be challenging in an intermittent system, but systems that
+allows dynamic checkpointing~\cite{hibernus}, or dynamic task formation~\cite{Dino} become
+even more difficult to program. Particularly in systems like ~\cite{Dino} where the
+programmer is responsible for placing task boundaries, there is a risk that programmers
+will either overuse task boundaries in an effort to preserve any given thread's state, or
+underprovision task boundaries such that progress is never made because of context
+switching that dynamically expands the number of instructions between task boundaries.
+
+While prior work has stopped short of suggesting an operating system for intermittently
+powered devices, a number of different works touch on the need for these devices to
+perform multiple tasks in the same workload. Buettner, Greenstein and Wetherall motivate
+their runtime, Dewdrop, based on difference in energy requirements for the tasks that a
+CRFID needs to perform~\cite{dewdrop}. The energy aware memory mapping in ~\cite{Aware}
+discusses the challenge of handling nondeterministic interrupts received by intermittent
+sensing platforms when a peripheral must be serviced immediately and control flow deviates
+from single threaded execution. From a hardware standpoint, the federated energy scheme in
+~\cite{ufop} provides a platform for removing the energy dependence between an MCU and its
+peripherals. Federated energy also simplifies a programmer's reasoning about reading from
+a single peripheral. However, federating energy does not help a programmer describe how to
+handle {\em multiple} peripherals. Consider a listing in
+~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
+of the hardware behavior a programmer would have to employ to, say, read one sensor if it
+is available but read from and proceed to process data from a different sensor if it is not.
+
+% Doesn't compile, commenting out
+%\begin{figure}[h]
+%  \centering
+%  \includegraphics[width=0.99\columnwidth]{figs/listing}
+%  \caption{{\bf Example program for sensing and sending data}}
+%  \label{fig:copy}
+%\end{figure}
+
+The wireless sensor networks community has explored multiple approaches for handling
+multitasking on resource limited devices. The designs tend to provide a combination of
+timer and event driven context switching as well as a simple programming interface.
+TinyOS uses a simple FIFO scheduler to handle event based context switches~\cite{tiny}.
+Nano-RK implements priority based preemptive multitasking where task priorities are
+statically defined~\cite{nano}. Contiki uses proto-theads, stackless program threads, to
+offer a model between nano-RK and tinyOS. The Contiki implements "cooperative
+multithreading" such that events can trigger a context switch, but multithreading may also
+be timer based~\cite{contiki}. The goal of multitasking interfaces in wireless sensor
+networks is to provide a tradeoff between network lifetime and quality of service. This
+differs from the intermittent programming model where the challenge is not system
+lifetime, but system functionality based on energy availability.
 
@@
 
+% Doesn't compile, commenting out
+%\begin{figure}[ht]
+%  \centering
+%  \includegraphics[width=0.9\columnwidth]{figs/tapir}
+%\end{figure}
+
 \balance

### 2017-05-11 paper.tex
Added pictures back in, figs/listing.PNG->figs/listing.png so it compiled
Added mutex section

@@
 
-Example Citation\cite{Chain, Mementos, Aware, Dino}!
-
-%Sample of adding a picture across the top of a split page...
-
-%\begin{figure*}
-%\centering
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
-%\caption{Capybara power system overview}\label{label-a}
-%\end{minipage}\hfill
-%\begin{minipage}[b]{0.49\textwidth}
-%  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
-%\caption{Effect of different power modes}\label{label-b}
-%\end{minipage}
-%\end{figure*}
-
-More text after the picture to show how that looks...
-
 \section{Related Work} 
@@
 % Doesn't compile, commenting out
-%\begin{figure}[h]
-%  \centering
-%  \includegraphics[width=0.99\columnwidth]{figs/listing}
-%  \caption{{\bf Example program for sensing and sending data}}
-%  \label{fig:copy}
-%\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.99\columnwidth]{figs/listing}
+  \caption{{\bf Example program for sensing and sending data}}
+  \label{fig:copy}
+\end{figure}
 
@@
 
-\section{What we did} % TODO change title
+\section{System Implementation } % TODO change title
 Mutexes - Emily\\
@@
 \texttt{TRANSITION\_TO} to also include thread creation.
+
+\subsection{Locking Primitive} 
+To improve a programmer's control of shared data between threads, and to reduce
+the complexity of accessing data that may be touched by many different tasks we
+implement a simple mutex object that can be added to a channel. The mutex is
+designed to be used in a standard \texttt{channel}, not a double buffered
+\texttt{self\_channel}, so it must be able to tolerate a failure after a
+read/write operation (i.e. test and set). To achieve idempotence, we add a
+\texttt{thread\_id} to the \texttt{free} bit inside of each mutex. When
+attempting to acquire a lock, a thread checks if its thread id is equal to the
+lock holder, then it checks if the lock is free. If either condition is true,
+the thread successfully acquires the lock, otherwise the \texttt{mutex\_lock}
+function fails. Upon successfully acquiring the lock, a thread writes its id to
+the mutex's holder id, and changes the \texttt{free} bit to \texttt{status =
+held}. This scheme guarantees correctness because if a failure occurs after the
+mutex is set but before the task commits, the task will read out the id of the
+mutex holder before checking if it is free. Upon matchint its id with the holder
+id, the task will proceed regardless of the fact that \texttt{status = held}.
+%Maybe add a figure here to make the sequence of read/write/failure clearer... 
+Further, a task must finsih executing before switching to another thread, so
+there is no possiblity for an external thread to see a value that was not
+committed by another task.
+
+It is important to note that the mutex object can be modified idempotently, but
+accesses to the data in the same channel as the mutex must adhere to channel
+exclusion. In our example application, the mutex is contained within a user
+defined \texttt{shared\_channel} structure that is analogous to a
+\texttt{multicast\_channel} but abandons the naming convention that "restricts"
+the channel to a static subset of tasks. Any task may read {\em or} or write to
+the \texttt{shared\_channel}, but performing both would break correctness. In
+the future, a compiler could be used to guarantee that a
+\texttt{shared\_channel} is not read and written by the same task, but at
+present we believe that the \texttt{mutex} and \texttt{shared\_channel}
+interface is easier for programmers to use than \texttt{multicast\_channels}, or
+a sepcialized double buffered channel when passing data between threads. It
+maintains familiar semantics for programmers and does not incur the memory
+overhead of double buffering.
 

### 2017-05-11 paper.tex
scheduling, SIMD vs. MIMD talk in background

--- 2017-05-11 paper.tex
@@
 
-Example Citation\cite{Chain, Mementos, Aware, Dino}!
+%Example Citation\cite{Chain, Mementos, Aware, Dino}!
 
@@
 
-More text after the picture to show how that looks...
+%More text after the picture to show how that looks...
 
@@
 Mutexes - Emily\\
-Round Robin (alternatives, on reboot)\\
-SIMD vs MIMD\\
+
+TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
+inserts a jump back to the scheduler task, where the next task in the list is
+the one to be executed. An alternate scheduling model that was considered was
+a context switch on reboot instead of at the end of each task. The main drawback
+of using a reboot-based scheduler is that, because power failure is the common
+case, there is a high chance of reducing or halting forward progress altogether
+because one thread holds a lock that all of the other threads need but cannot
+complete. For that reason the design specifies end of task, so each task has
+an opportunity to use a shared resource and release any locks before a power-
+intensive context switch occurs. Additionally, a round robin scheduler was 
+chosen for this design because it is simple to implement and uses only one index counter,
+very little power or memory overhead as compared to other scheduling algorithms. It is 
+critical for multithreading in Chain to not add significant power and memory
+overhead in scheduling over conventional Chain so that the workload can
+make as much forward progress as possible on one charge cycle.
+
+When designing threading in Chain, the shared memory model is a main indicator 
+of the threaded execution model. The two most probable options when designing 
+TAPIR were a SIMD execution model of \textit{Thread Pools}, or a MIMD 
+traditional multiprogramming model. The Thread Pool design makes use of task 
+code duplication by explicitly forking a number of chain programs, all of which 
+run simultaneously on the scheduler on arrayed data. This option is useful 
+multiple data streams from I/O in embedded applications that need to be 
+processed before new data overrides the I/O device. Additionally, the Thread 
+Pool model makes sharing memory more regular by using multiple copies of the 
+same thread. However, this option is not as general purpose. The TAPIR design 
+ultimately settled on a multiprogramming model, because this allows for the 
+programmer to explictly program unique chain programs with shared or discrete 
+data structures on the same system. However, this execution model requires a
+more structured locking mechanism to avoid RMW during execution.
+
 SIMD w/ channels (you don't get your own registers...etc -

### 2017-05-11 paper.tex

--- 2017-05-11 paper.tex
@@
 
-% Doesn't compile, commenting out
-%\begin{figure}[ht]
-%  \centering
-%  \includegraphics[width=0.9\columnwidth]{figs/tapir}
-%\end{figure}
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=0.9\columnwidth]{figs/tapir}
+\end{figure}
 

### 2017-05-11 paper.tex
Added more application description, fixed broken refs

--- 2017-05-11 paper.tex
@@
 \usepackage{listings}
+\usepackage{url}
 \usepackage{xspace}
@@
   \label{sec:intro}
-%A/N: I don't think that the paragraphs in this section are in a great order, but I am having a hard time making it flow better.
-%Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
+%A/N: I don't think that the paragraphs in this section are in a great order,
+%but I am having a hard time making it flow better.  Also I'm absolutely sure
+%each of these paragraphs can definitely be bulked up more for length. Please
+%help thank you!!!
 
-The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
-New technologies in both of these areas has led to new
-\textit{energy-harvesting computing devices} that operates only with power it has
-collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.
+The increase in popularity of Internet of Things (IoT) has led to research focus
+in wireless technologies and low-power computing.  New technologies in both of
+these areas has led to new \textit{energy-harvesting computing devices} that
+operates only with power it has collected from the surrounding environment
+\cite{Chain}. These devices have use cases in medical, IoT, and research
+applications.
 
-Commonly, an energy-harvesting devices charges and collects power in a storage capacitor until the capacitor reaches a sufficient level of charge.
-Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
-This leads to the \textit{intermittent execution model} where regular shut downs and reboots to recharge are the common case rather than the rare case in
-a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave
-irregular behavior on an intermittent device and a new programming model is needed.
+Commonly, an energy-harvesting devices charges and collects power in a storage
+capacitor until the capacitor reaches a sufficient level of charge.  Once the
+capacitor is charged, the device begins its operation phase and will run until
+the capacitor is drained.  This leads to the \textit{intermittent execution
+model} where regular shut downs and reboots to recharge are the common case
+rather than the rare case in a traditional execution model. Because the
+intermittent execution model includes power failures and reboots, continuous
+execution programs behave irregular behavior on an intermittent device and a new
+programming model is needed.
 
-Our multithreaded model is build on two existing programming models, DINO and Chain.
-An initial intermittent programming model DINO breaks intermittent programs into tasks, which creates automatic checkpointing locations
-during exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks both volatile and nonvolatile memory.
-The DINO execution model executes all tasks transactionally, leading to atomic task semantics and does not permit consistency violations in
-nonvolatile state \cite{Dino}. Chain, a more recent iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses task boundaries
-called channels to store most used data in nonvolatile memory \cite{Chain}. At compile time, Chain allocates the necessary channels between tasks
-by parsing channel read and write statements in order to match up aligning tasks. All inputs and outputs to each task, as well as necessary data
-structures, are stored in a channel for a respective task. A channel may only be modified by one task, and only one task may run at a time,
-which mitigates time of check, time of use bugs (TOCTOU) in intermittence.
-The one exception is a self-channel for recursive code, which is double buffered.
-Keeping all modifiable data in nonvolatile memory that is not open to easy rewrite makes these tasks idempotent across reboots.
-Because of continuous reboots, many operations necessarily will be executed more than once,
-and correctness is compromised if those operations are not guaranteed to have the same outcome by the programming model after each reboot.
+Our multithreaded model is build on two existing programming models, DINO and
+Chain.  An initial intermittent programming model DINO breaks intermittent
+programs into tasks, which creates automatic checkpointing locations during
+exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks
+both volatile and nonvolatile memory.  The DINO execution model executes all
+tasks transactionally, leading to atomic task semantics and does not permit
+consistency violations in nonvolatile state \cite{Dino}. Chain, a more recent
+iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses
+task boundaries called channels to store most used data in nonvolatile memory
+\cite{Chain}. At compile time, Chain allocates the necessary channels between
+tasks by parsing channel read and write statements in order to match up aligning
+tasks. All inputs and outputs to each task, as well as necessary data
+structures, are stored in a channel for a respective task. A channel may only be
+modified by one task, and only one task may run at a time, which mitigates time
+of check, time of use bugs (TOCTOU) in intermittence.  The one exception is a
+self-channel for recursive code, which is double buffered.  Keeping all
+modifiable data in nonvolatile memory that is not open to easy rewrite makes
+these tasks idempotent across reboots.  Because of continuous reboots, many
+operations necessarily will be executed more than once, and correctness is
+compromised if those operations are not guaranteed to have the same outcome by
+the programming model after each reboot.
 
-A main issue in intermittent computing along with idempotence is the Read-Modify-Write (RMW) problem.
-Because volatile memory is cleared on each reboot, it is possible that in between reading and modifying or modifying and writing the data had changed,
-leading to incorrect answers and an inconsistent state. In the Chain programming model, this problem is solved by creating a new task to house the
-problematic operations. Because tasks are blocked by channels that can only be written by one task and are saved to nonvolatile memory, RMW is avoided.
-However, in multiprogramming the problem manifests in another way; if the scheduler switches tasks on a reboot, the currently running thread could have
-been in the middle of its critical section, making it difficult for other threads to make progress if context switches always occur after a reboot.
-Because of this we have decided on a round-robin scheduler with context switches at the end of a task, to guarantee forward progress while knowing
-which tasks have run to completion.
+A main issue in intermittent computing along with idempotence is the
+Read-Modify-Write (RMW) problem.  Because volatile memory is cleared on each
+reboot, it is possible that in between reading and modifying or modifying and
+writing the data had changed, leading to incorrect answers and an inconsistent
+state. In the Chain programming model, this problem is solved by creating a new
+task to house the problematic operations. Because tasks are blocked by channels
+that can only be written by one task and are saved to nonvolatile memory, RMW is
+avoided.  However, in multiprogramming the problem manifests in another way; if
+the scheduler switches tasks on a reboot, the currently running thread could
+have been in the middle of its critical section, making it difficult for other
+threads to make progress if context switches always occur after a reboot.
+Because of this we have decided on a round-robin scheduler with context switches
+at the end of a task, to guarantee forward progress while knowing which tasks
+have run to completion.
 
-In the current Chain programming model, writing code that completes multiple objectives correctly is now difficult.
-Current chain programs are small and only have one objective. Additionally, there is no support for advanced control flow.
-Multiple control threads may exist, but in the current model there is no support for them and programmability is difficult and often incorrect.
-Supporting multiple control threads in Chain also necessitates a linear increase in a limited memory resource, because all large shared data structures
-must be copied into each necessary channel, and also makes updating the structures a time consuming, high-overhead task.
-In order to increase functionality of chain and to mitigate these shortcomings, we add familiar threading semantics to Chain in order to express
-multi-threaded control flow operation. In order to do this we create a lightweight scheduler task that becomes part of the inter-task checkpointing
-operation and provide a method of sharing large data structures between tasks with a locking feature. This method is low-memory and low-execution
-overhead, which can then maintain the Chain programming semantics at the user level still successfully complete work on an intermittent device.
+In the current Chain programming model, writing code that completes multiple
+objectives correctly is now difficult.  Current chain programs are small and
+only have one objective. Additionally, there is no support for advanced control
+flow.  Multiple control threads may exist, but in the current model there is no
+support for them and programmability is difficult and often incorrect.
+Supporting multiple control threads in Chain also necessitates a linear increase
+in a limited memory resource, because all large shared data structures must be
+copied into each necessary channel, and also makes updating the structures a
+time consuming, high-overhead task.  In order to increase functionality of chain
+and to mitigate these shortcomings, we add familiar threading semantics to Chain
+in order to express multi-threaded control flow operation. In order to do this
+we create a lightweight scheduler task that becomes part of the inter-task
+checkpointing operation and provide a method of sharing large data structures
+between tasks with a locking feature. This method is low-memory and
+low-execution overhead, which can then maintain the Chain programming semantics
+at the user level still successfully complete work on an intermittent device.
 
@@
 handle {\em multiple} peripherals. Consider a listing in
-~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
+~\cite{ufop} reporoduced in Figure ~\ref{fig:copy}, and imagine the level of understanding
 of the hardware behavior a programmer would have to employ to, say, read one sensor if it
@@
 
-\section{Implementation}
-Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
-Mutexes (Simple, meant for pseudo channels) - Emily\\
-Interrupts (mucked with IE flags) - Neil\\
-% Interrupt prologue
-% Interrupt setup
+%\section{Implementation}
+%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
+%Mutexes (Simple, meant for pseudo channels) - Emily\\
+%Interrupts (mucked with IE flags) - Neil\\
+%% Interrupt prologue
+%% Interrupt setup
 
-\section{Results} - Emily
-No really, there will be things here!
-Applications/measurements\\
-Blinker (x2)\\
-RSA+cukoo\\
-No verification\\
+\section{Evaluation}
+To test our multithreading library and mutex implementation, we instrumented
+several existing Chain applications that do not share any tasks or channels to
+run concurrently. The programs were run on either an msp430fr5969 launchpad or
+on a Capybara sensing node, a custom intermittent development board with an
+msp430fr5949 MCU. Reboot tests were carried out strictly on the Capybara board
+because was simpler to guarantee the isolation of the MCU from the programming
+fet than on the launchpad. We did not seekt to fully verify our implementation
+of multi-threading, but we use our experience in application development to
+support our claim that the threads behave correctly, and as a programmer would
+expect. 
+\subsection{Applications}
+\subsubsection{Blinker (B)}
+This application was built on top of the app-blinker-chain
+example~\cite{blinker} and modified to remove the delays and LED activations, so
+that it just passes data through three tasks until a maximum count is reached,
+and then the thread terminates. This app was primarily used to demonstrate the
+initial functionality of the system, and then to test mutexes. In both cases a
+combination of \texttt{LOG} statements and final output values were used to
+confirm the correct operation of the system. 
+\subsubsection{Cuckoo-RSA (CRSA)}
+This application is a combination of a cuckoo filter that inserts and queries 64
+values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
+using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
+combination where merging the original applications did not produce the same
+final result as running the applications separately. Several bytes at the end of
+a crucial array for RSA were always overwritten when the applications were
+merged, and we were unable to fix the problem. Unit tests running the RSA
+application in single threaded mode through our scheduler and using our
+multi-threading library produce the correct result, but any combination of RSA
+and cuckoo filtering produces the incorrect result. However, RSA does run to
+completion, so we believe the results from C-RSA are still useful. 
+\subsubsection{Cuckoo-Bitcount (CB)}
+This application is a combination of the cuckoo filter described above and a
+application that uses different algorithms to count the number of bits in an
+array of integers~\cite{bitcount}. CB is a correctly working combination of two
+applications that produces the same output as separate executions of the
+unmodified versions of these applications. It expands the reliability of our
+multi-threading library beyond simple applications like Blinker.
+\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
+CBB combines the cuckoo filter and bit count applications with two threads of
+the blinker application. This results in four active threads at once, which is
+the maximum number of threads our prototype implementation allows. We confirmed
+that all four threads are fairly serviced by the round robin scheduler and they
+correctly start and end.
+
 Memory overhead (read off compilation)\\
@@
 \section{Future Work}
-Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks.
-When forking a new task, a channel may be duplicated, which linearly increases the amount of memory necessary for channels with the number of currently running programs.
-It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how
-many are to be running simultaneously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
-In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
-Currently channels are allocated and named at compile time by combining the names of the input and output channel tasks.
-However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
-This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.
+Future work is focused on saving memory by using channel and code reuse
+optimizations when forking new chain programs from current tasks.  When forking
+a new task, a channel may be duplicated, which linearly increases the amount of
+memory necessary for channels with the number of currently running programs.  It
+is possible to reuse channels between tasks if the number of iterations of that
+task are known at compile time and it can be estimated how many are to be
+running simultaneously. A register allocation graph coloring scheme can then be
+repurposed for reusing channels allocated at compile time.  In either case the
+current channel naming scheme must be changed to support channel duplication,
+ether real or simulated.  Currently channels are allocated and named at compile
+time by combining the names of the input and output channel tasks.  However, if
+tasks can be duplicated this naming scheme must be altered to support multiple
+channels of the same name and potentially multiple channels of the same name.
+This can be done with a vector clock timestamp or a simple counter stored in
+nonvolatile memory to create unique channel names.
 
-Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
-To do this the concept of intermittent tasks mutexes for shared resources and data structures would be introduced.
-The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
-In order to allow all tasks access to a data structure without breaking all chain semantics, a modified version of a module channel would be used in order
-for multiple tasks to potentially modify one channel.
-However, even with that change it is challenging to define sharing between threads in an intermittent context and to maintain chain semantics while implementing
-shared data structures while still maintaining correctness.
-Because channel sizes are constrained at compile time currently, it is also necessary to know at compile time all data structures, shared or private,
-that are going to be used during execution of a chain program or programs.
-It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory
-is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.
+Additionally, it is possible to save memory by sharing arrays between tasks
+without the necessity of copying the array in its entirety in each channel.  To
+do this the concept of intermittent tasks mutexes for shared resources and data
+structures would be introduced.  The structures are protected by a lock that
+contains the thread ID of the task that currently holds the lock.  In order to
+allow all tasks access to a data structure without breaking all chain semantics,
+a modified version of a module channel would be used in order for multiple tasks
+to potentially modify one channel.  However, even with that change it is
+challenging to define sharing between threads in an intermittent context and to
+maintain chain semantics while implementing shared data structures while still
+maintaining correctness.  Because channel sizes are constrained at compile time
+currently, it is also necessary to know at compile time all data structures,
+shared or private, that are going to be used during execution of a chain program
+or programs.  It is a challenge for future work in this field to constrain
+memory for channels at compile time and for the compiler to know with certainty
+how much memory is necessary to allocate. Depending on the workload, the amount
+of forked threads would need to be constrained such that threads don't overflow
+nonvolatile memory at runtime.
 
-Future work also includes optimizations to reduce overhead, such as converting the free indices array to a free indices bitmap in order to reduce the amount of
-baseline memory that the scheduler task uses.
-Re-assignation and re-use of channels at runtime would increase overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
-If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.
+Future work also includes optimizations to reduce overhead, such as converting
+the free indices array to a free indices bitmap in order to reduce the amount of
+baseline memory that the scheduler task uses.  Re-assignation and re-use of
+channels at runtime would increase overhead, but it is thought that the memory
+saved by this procedure is worth the trade-off of increased overhead.  If
+re-assignation can be determined at compile-time, the overhead during runtime
+will still remain low.
 

### 2017-05-11 paper.tex
Added section on scheduler data structures Added style to make code listings nicer Trailing whitespace fixes

--- 2017-05-11 paper.tex
@@
 \newcommand{\chain}{Chain\xspace}
+
+\usepackage{color}
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.95}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\footnotesize,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    keepspaces=true,
+    numbers=none,
+    numbersep=5pt,
+    showspaces=false,
+    showstringspaces=false,
+    showtabs=false,
+    tabsize=4
+    }
+    \lstset{style=mystyle}
+
 \numberofauthors{3}
@@
 
+%Example Citation\cite{Chain, Mementos, Aware, Dino}!
+
+%Sample of adding a picture across the top of a split page...
+
+%\begin{figure*}
+%\centering
+%\begin{minipage}[b]{0.49\textwidth}
+%  \includegraphics[width=0.95\textwidth,center]{capybara-all.png}
+%\caption{Capybara power system overview}\label{label-a}
+%\end{minipage}\hfill
+%\begin{minipage}[b]{0.49\textwidth}
+%  \includegraphics[width=0.95\textwidth,center]{power_modes.pdf}
+%\caption{Effect of different power modes}\label{label-b}
+%\end{minipage}
+%\end{figure*}
+
+%More text after the picture to show how that looks...
+
 \section{Related Work}
@@
 
-% Doesn't compile, commenting out
 \begin{figure}[h]
@@
 
-\section{System Implementation } % TODO change title
+\section{What we did} % TODO change title
 Mutexes - Emily\\
-
-TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
-inserts a jump back to the scheduler task, where the next task in the list is
-the one to be executed. An alternate scheduling model that was considered was
-a context switch on reboot instead of at the end of each task. The main drawback
-of using a reboot-based scheduler is that, because power failure is the common
-case, there is a high chance of reducing or halting forward progress altogether
-because one thread holds a lock that all of the other threads need but cannot
-complete. For that reason the design specifies end of task, so each task has
-an opportunity to use a shared resource and release any locks before a power-
-intensive context switch occurs. Additionally, a round robin scheduler was 
-chosen for this design because it is simple to implement and uses only one index counter,
-very little power or memory overhead as compared to other scheduling algorithms. It is 
-critical for multithreading in Chain to not add significant power and memory
-overhead in scheduling over conventional Chain so that the workload can
-make as much forward progress as possible on one charge cycle.
-
-When designing threading in Chain, the shared memory model is a main indicator 
-of the threaded execution model. The two most probable options when designing 
-TAPIR were a SIMD execution model of \textit{Thread Pools}, or a MIMD 
-traditional multiprogramming model. The Thread Pool design makes use of task 
-code duplication by explicitly forking a number of chain programs, all of which 
-run simultaneously on the scheduler on arrayed data. This option is useful 
-multiple data streams from I/O in embedded applications that need to be 
-processed before new data overrides the I/O device. Additionally, the Thread 
-Pool model makes sharing memory more regular by using multiple copies of the 
-same thread. However, this option is not as general purpose. The TAPIR design 
-ultimately settled on a multiprogramming model, because this allows for the 
-programmer to explictly program unique chain programs with shared or discrete 
-data structures on the same system. However, this execution model requires a
-more structured locking mechanism to avoid RMW during execution.
-
+Round Robin (alternatives, on reboot)\\
+SIMD vs MIMD\\
 SIMD w/ channels (you don't get your own registers...etc -
@@
 \texttt{TRANSITION\_TO} to also include thread creation.
-
-\subsection{Locking Primitive} 
-To improve a programmer's control of shared data between threads, and to reduce
-the complexity of accessing data that may be touched by many different tasks we
-implement a simple mutex object that can be added to a channel. The mutex is
-designed to be used in a standard \texttt{channel}, not a double buffered
-\texttt{self\_channel}, so it must be able to tolerate a failure after a
-read/write operation (i.e. test and set). To achieve idempotence, we add a
-\texttt{thread\_id} to the \texttt{free} bit inside of each mutex. When
-attempting to acquire a lock, a thread checks if its thread id is equal to the
-lock holder, then it checks if the lock is free. If either condition is true,
-the thread successfully acquires the lock, otherwise the \texttt{mutex\_lock}
-function fails. Upon successfully acquiring the lock, a thread writes its id to
-the mutex's holder id, and changes the \texttt{free} bit to \texttt{status =
-held}. This scheme guarantees correctness because if a failure occurs after the
-mutex is set but before the task commits, the task will read out the id of the
-mutex holder before checking if it is free. Upon matchint its id with the holder
-id, the task will proceed regardless of the fact that \texttt{status = held}.
-%Maybe add a figure here to make the sequence of read/write/failure clearer... 
-Further, a task must finsih executing before switching to another thread, so
-there is no possiblity for an external thread to see a value that was not
-committed by another task.
-
-It is important to note that the mutex object can be modified idempotently, but
-accesses to the data in the same channel as the mutex must adhere to channel
-exclusion. In our example application, the mutex is contained within a user
-defined \texttt{shared\_channel} structure that is analogous to a
-\texttt{multicast\_channel} but abandons the naming convention that "restricts"
-the channel to a static subset of tasks. Any task may read {\em or} or write to
-the \texttt{shared\_channel}, but performing both would break correctness. In
-the future, a compiler could be used to guarantee that a
-\texttt{shared\_channel} is not read and written by the same task, but at
-present we believe that the \texttt{mutex} and \texttt{shared\_channel}
-interface is easier for programmers to use than \texttt{multicast\_channels}, or
-a sepcialized double buffered channel when passing data between threads. It
-maintains familiar semantics for programmers and does not incur the memory
-overhead of double buffering.
 
@@
 \begin{minipage}[b]{1.0\textwidth}
-\begin{lstlisting}
+\begin{lstlisting}[language=C]
     /** Enable interrupts if int_setup_complete() has been called
@@
 \section{Implementation}
-Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
+\subsection{Scheduler}
+\begin{figure*}
+\begin{minipage}[b]{.48\textwidth}
+\begin{lstlisting}[language=C]
+    typedef struct thread_t {
+        unsigned thread_id;
+        context_t context;
+    } thread_t;
+
+    typedef struct thread_state_t {
+        thread_t thread;
+        unsigned active;
+    } thread_state_t;
+
+    // Array of running threads
+    thread_state_t threads[MAX_NUM_THREADS];
+
+    // List of free indicies in threads[]
+    unsigned free_indicies[MAX_NUM_THREADS];
+    unsigned free_indicies_size;
+
+    // Index of current running thread in threads[]
+    // Stored in self channel
+    unsigned current;
+\end{lstlisting}
+\caption{Scheduler Data Structures}\label{schDataStruct}
+\end{minipage}\hfill
+\end{figure*}
+
+The data structures for our scheduler are listed in figure \ref{schDataStruct}.
+Like the rest of chain, memory for data structures is allocated at compile
+time. Thus, the limit on the maximum number of threads allowed in a system is
+fixed arbitrarily - not ideal, but there isn't a better option while operating
+inside chain. The \texttt{threads} array is stored in all-to-one channel, from
+every task to the scheduler task. The \texttt{free\_indicies} array is stored
+in a one-to-all channel, from the scheduler task to every task.
+
+Neither all-to-one nor one-to-all channels fit into standard \chain channel
+semantics, although they resemble multicast-in and multicast-out channels
+described in the original \chain paper. We eschew the notion that channels
+have fixed start and end points and instead treat the channel as a general
+location in non-volatile memory that can be read from and written to by
+arbitrary tasks. However, accesses to these channels are encapsulated by our
+library, so the programmer has no need to think about idempotency violations.
+
+With these multicast channels, the scheduler task can act as a buffer between
+multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
+\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
+\texttt{free\_indicies} through calls to thread create, maintaining
+idempotency. Inside the scheduler task, we read and write \texttt{current},
+so \texttt{current} is double buffered. The only issue is with
+\texttt{thread\_init} since it is called from user-defined tasks and
+initializes \texttt{current} - the same task that called \texttt{thread\_init}
+could call \texttt{thread\_create}, and the \texttt{thread\_create} call
+wouldn't see the update to the variable. Since \texttt{current} is always
+initialized to the same value (0), we opt for manually swapping the buffer on
+\texttt{current} so that other threads can see the update.
+
+We keep an unsigned integer, \texttt{curr\_free\_index}, in volatile memory,
+zeroed on every reboot, that acts as a current pointer in
+\texttt{free\_indicies[]}. Every time the scheduler task runs,
+\texttt{free\_indicies[]} is updated with the state of \texttt{threads[]},
+based on whether or not the active bit is set for a given index. When
+\texttt{thread\_create} is called, if \texttt{curr\_free\_index} is greater
+than \texttt{free\_indicies\_size}, then \texttt{thread\_create} will fail in
+POSIX fashion - the system does not have enough resources to support another
+thread. Otherwise, \texttt{threads[free\_indicies[curr\_free\_index]]} is set
+to the new thread. Upon power failure, \texttt{curr\_free\_index} is reset to
+zero, so as long as the rest of the operations in the task are idempotent, so
+will \texttt{thread\_create} calls - the same locations in \texttt{theads[]}
+will be overwritten.
+
 Mutexes (Simple, meant for pseudo channels) - Emily\\

### 2017-05-11 paper.tex

--- 2017-05-11 paper.tex
@@
 
+\section{Implementation}
+\subsection{Scheduler}
+\begin{figure*}
+\begin{minipage}[b]{.48\textwidth}
+\begin{lstlisting}[language=C]
+    typedef struct thread_t {
+        unsigned thread_id;
+        context_t context;
+    } thread_t;
+
+    typedef struct thread_state_t {
+        thread_t thread;
+        unsigned active;
+    } thread_state_t;
+
+    // Array of running threads
+    thread_state_t threads[MAX_NUM_THREADS];
+
+    // List of free indicies in threads[]
+    unsigned free_indicies[MAX_NUM_THREADS];
+    unsigned free_indicies_size;
+
+    // Index of current running thread in threads[]
+    // Stored in self channel
+    unsigned current;
+\end{lstlisting}
+\caption{Scheduler Data Structures}\label{schDataStruct}
+\end{minipage}\hfill
+\end{figure*}
+
+The data structures for our scheduler are listed in figure \ref{schDataStruct}.
+Like the rest of chain, memory for data structures is allocated at compile
+time. Thus, the limit on the maximum number of threads allowed in a system is
+fixed arbitrarily - not ideal, but there isn't a better option while operating
+inside chain. The \texttt{threads} array is stored in all-to-one channel, from
+every task to the scheduler task. The \texttt{free\_indicies} array is stored
+in a one-to-all channel, from the scheduler task to every task.
+
+Neither all-to-one nor one-to-all channels fit into standard \chain channel
+semantics, although they resemble multicast-in and multicast-out channels
+described in the original \chain paper. We eschew the notion that channels
+have fixed start and end points and instead treat the channel as a general
+location in non-volatile memory that can be read from and written to by
+arbitrary tasks. However, accesses to these channels are encapsulated by our
+library, so the programmer has no need to think about idempotency violations.
+
+With these multicast channels, the scheduler task can act as a buffer between
+multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
+\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
+\texttt{free\_indicies} through calls to thread create, maintaining
+idempotency. Inside the scheduler task, we read and write \texttt{current},
+so \texttt{current} is double buffered. The only issue is with
+\texttt{thread\_init} since it is called from user-defined tasks and
+initializes \texttt{current} - the same task that called \texttt{thread\_init}
+could call \texttt{thread\_create}, and the \texttt{thread\_create} call
+wouldn't see the update to the variable. Since \texttt{current} is always
+initialized to the same value (0), we opt for manually swapping the buffer on
+\texttt{current} so that other threads can see the update.
+
+We keep an unsigned integer, \texttt{curr\_free\_index}, in volatile memory,
+zeroed on every reboot, that acts as a current pointer in
+\texttt{free\_indicies[]}. Every time the scheduler task runs,
+\texttt{free\_indicies[]} is updated with the state of \texttt{threads[]},
+based on whether or not the active bit is set for a given index. When
+\texttt{thread\_create} is called, if \texttt{curr\_free\_index} is greater
+than \texttt{free\_indicies\_size}, then \texttt{thread\_create} will fail in
+POSIX fashion - the system does not have enough resources to support another
+thread. Otherwise, \texttt{threads[free\_indicies[curr\_free\_index]]} is set
+to the new thread. Upon power failure, \texttt{curr\_free\_index} is reset to
+zero, so as long as the rest of the operations in the task are idempotent, so
+will \texttt{thread\_create} calls - the same locations in \texttt{theads[]}
+will be overwritten.
+
+% TODO picture of array?
+
+Mutexes (Simple, meant for pseudo channels) - Emily\\
+Interrupts (mucked with IE flags) - Neil\\
+% Interrupt prologue
+% Interrupt setup
 %\section{Implementation}
@@
 expect.
+
 \subsection{Applications}

### 2017-05-11 paper.tex
more stuff about SIMD

--- 2017-05-11 paper.tex
@@
 \usepackage{listings}
-\usepackage{url}
 \usepackage{xspace}
@@
   \label{sec:intro}
-%A/N: I don't think that the paragraphs in this section are in a great order,
-%but I am having a hard time making it flow better.  Also I'm absolutely sure
-%each of these paragraphs can definitely be bulked up more for length. Please
-%help thank you!!!
+%A/N: I don't think that the paragraphs in this section are in a great order, but I am having a hard time making it flow better.
+%Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
 
-The increase in popularity of Internet of Things (IoT) has led to research focus
-in wireless technologies and low-power computing.  New technologies in both of
-these areas has led to new \textit{energy-harvesting computing devices} that
-operates only with power it has collected from the surrounding environment
-\cite{Chain}. These devices have use cases in medical, IoT, and research
-applications.
+The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
+New technologies in both of these areas has led to new
+\textit{energy-harvesting computing devices} that operates only with power it has
+collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.
 
-Commonly, an energy-harvesting devices charges and collects power in a storage
-capacitor until the capacitor reaches a sufficient level of charge.  Once the
-capacitor is charged, the device begins its operation phase and will run until
-the capacitor is drained.  This leads to the \textit{intermittent execution
-model} where regular shut downs and reboots to recharge are the common case
-rather than the rare case in a traditional execution model. Because the
-intermittent execution model includes power failures and reboots, continuous
-execution programs behave irregular behavior on an intermittent device and a new
-programming model is needed.
+Commonly, an energy-harvesting devices charges and collects power in a storage capacitor until the capacitor reaches a sufficient level of charge.
+Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
+This leads to the \textit{intermittent execution model} where regular shut downs and reboots to recharge are the common case rather than the rare case in
+a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave
+irregular behavior on an intermittent device and a new programming model is needed.
 
-Our multithreaded model is build on two existing programming models, DINO and
-Chain.  An initial intermittent programming model DINO breaks intermittent
-programs into tasks, which creates automatic checkpointing locations during
-exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks
-both volatile and nonvolatile memory.  The DINO execution model executes all
-tasks transactionally, leading to atomic task semantics and does not permit
-consistency violations in nonvolatile state \cite{Dino}. Chain, a more recent
-iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses
-task boundaries called channels to store most used data in nonvolatile memory
-\cite{Chain}. At compile time, Chain allocates the necessary channels between
-tasks by parsing channel read and write statements in order to match up aligning
-tasks. All inputs and outputs to each task, as well as necessary data
-structures, are stored in a channel for a respective task. A channel may only be
-modified by one task, and only one task may run at a time, which mitigates time
-of check, time of use bugs (TOCTOU) in intermittence.  The one exception is a
-self-channel for recursive code, which is double buffered.  Keeping all
-modifiable data in nonvolatile memory that is not open to easy rewrite makes
-these tasks idempotent across reboots.  Because of continuous reboots, many
-operations necessarily will be executed more than once, and correctness is
-compromised if those operations are not guaranteed to have the same outcome by
-the programming model after each reboot.
+Our multithreaded model is build on two existing programming models, DINO and Chain.
+An initial intermittent programming model DINO breaks intermittent programs into tasks, which creates automatic checkpointing locations
+during exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks both volatile and nonvolatile memory.
+The DINO execution model executes all tasks transactionally, leading to atomic task semantics and does not permit consistency violations in
+nonvolatile state \cite{Dino}. Chain, a more recent iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses task boundaries
+called channels to store most used data in nonvolatile memory \cite{Chain}. At compile time, Chain allocates the necessary channels between tasks
+by parsing channel read and write statements in order to match up aligning tasks. All inputs and outputs to each task, as well as necessary data
+structures, are stored in a channel for a respective task. A channel may only be modified by one task, and only one task may run at a time,
+which mitigates time of check, time of use bugs (TOCTOU) in intermittence.
+The one exception is a self-channel for recursive code, which is double buffered.
+Keeping all modifiable data in nonvolatile memory that is not open to easy rewrite makes these tasks idempotent across reboots.
+Because of continuous reboots, many operations necessarily will be executed more than once,
+and correctness is compromised if those operations are not guaranteed to have the same outcome by the programming model after each reboot.
 
-A main issue in intermittent computing along with idempotence is the
-Read-Modify-Write (RMW) problem.  Because volatile memory is cleared on each
-reboot, it is possible that in between reading and modifying or modifying and
-writing the data had changed, leading to incorrect answers and an inconsistent
-state. In the Chain programming model, this problem is solved by creating a new
-task to house the problematic operations. Because tasks are blocked by channels
-that can only be written by one task and are saved to nonvolatile memory, RMW is
-avoided.  However, in multiprogramming the problem manifests in another way; if
-the scheduler switches tasks on a reboot, the currently running thread could
-have been in the middle of its critical section, making it difficult for other
-threads to make progress if context switches always occur after a reboot.
-Because of this we have decided on a round-robin scheduler with context switches
-at the end of a task, to guarantee forward progress while knowing which tasks
-have run to completion.
+A main issue in intermittent computing along with idempotence is the Read-Modify-Write (RMW) problem.
+Because volatile memory is cleared on each reboot, it is possible that in between reading and modifying or modifying and writing the data had changed,
+leading to incorrect answers and an inconsistent state. In the Chain programming model, this problem is solved by creating a new task to house the
+problematic operations. Because tasks are blocked by channels that can only be written by one task and are saved to nonvolatile memory, RMW is avoided.
+However, in multiprogramming the problem manifests in another way; if the scheduler switches tasks on a reboot, the currently running thread could have
+been in the middle of its critical section, making it difficult for other threads to make progress if context switches always occur after a reboot.
+Because of this we have decided on a round-robin scheduler with context switches at the end of a task, to guarantee forward progress while knowing
+which tasks have run to completion.
 
-In the current Chain programming model, writing code that completes multiple
-objectives correctly is now difficult.  Current chain programs are small and
-only have one objective. Additionally, there is no support for advanced control
-flow.  Multiple control threads may exist, but in the current model there is no
-support for them and programmability is difficult and often incorrect.
-Supporting multiple control threads in Chain also necessitates a linear increase
-in a limited memory resource, because all large shared data structures must be
-copied into each necessary channel, and also makes updating the structures a
-time consuming, high-overhead task.  In order to increase functionality of chain
-and to mitigate these shortcomings, we add familiar threading semantics to Chain
-in order to express multi-threaded control flow operation. In order to do this
-we create a lightweight scheduler task that becomes part of the inter-task
-checkpointing operation and provide a method of sharing large data structures
-between tasks with a locking feature. This method is low-memory and
-low-execution overhead, which can then maintain the Chain programming semantics
-at the user level still successfully complete work on an intermittent device.
+In the current Chain programming model, writing code that completes multiple objectives correctly is now difficult.
+Current chain programs are small and only have one objective. Additionally, there is no support for advanced control flow.
+Multiple control threads may exist, but in the current model there is no support for them and programmability is difficult and often incorrect.
+Supporting multiple control threads in Chain also necessitates a linear increase in a limited memory resource, because all large shared data structures
+must be copied into each necessary channel, and also makes updating the structures a time consuming, high-overhead task.
+In order to increase functionality of chain and to mitigate these shortcomings, we add familiar threading semantics to Chain in order to express
+multi-threaded control flow operation. In order to do this we create a lightweight scheduler task that becomes part of the inter-task checkpointing
+operation and provide a method of sharing large data structures between tasks with a locking feature. This method is low-memory and low-execution
+overhead, which can then maintain the Chain programming semantics at the user level still successfully complete work on an intermittent device.
 
@@
 handle {\em multiple} peripherals. Consider a listing in
-~\cite{ufop} reporoduced in Figure ~\ref{fig:copy}, and imagine the level of understanding
+~\cite{ufop} reporoduced in Figure ~\ref{copy}, and imagine the level of understanding
 of the hardware behavior a programmer would have to employ to, say, read one sensor if it
@@
 
-SIMD w/ channels (you don't get your own registers...etc -
-    just channels, since intermittent)\\
+SIMD execution on an intermittent system relies on nonvolatile memory instead
+of independent registers for each thread. In Chain's case the implementation
+would rely on channels for each thread running in the system. In the MIMD
+execution model channels could potentially be re-assigned to already allocated
+channels because the threads may end at interleaving times. However, in the
+SIMD model it is expected that each identical thread is at a comperable point
+in execution to the other running threads, and so each thread that may be
+run necessitates a new channel allocation. This could lead to an unavoidable
+hard limit on the number of threads that can run in an applicatioin which in
+turn limits potential applications that are useful on this intermittent device.
 
@@
 
-%\section{Implementation}
-%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
-%Mutexes (Simple, meant for pseudo channels) - Emily\\
-%Interrupts (mucked with IE flags) - Neil\\
-%% Interrupt prologue
-%% Interrupt setup
+\section{Implementation}
+Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
+Mutexes (Simple, meant for pseudo channels) - Emily\\
+Interrupts (mucked with IE flags) - Neil\\
+% Interrupt prologue
+% Interrupt setup
 
-\section{Evaluation}
-To test our multithreading library and mutex implementation, we instrumented
-several existing Chain applications that do not share any tasks or channels to
-run concurrently. The programs were run on either an msp430fr5969 launchpad or
-on a Capybara sensing node, a custom intermittent development board with an
-msp430fr5949 MCU. Reboot tests were carried out strictly on the Capybara board
-because was simpler to guarantee the isolation of the MCU from the programming
-fet than on the launchpad. We did not seekt to fully verify our implementation
-of multi-threading, but we use our experience in application development to
-support our claim that the threads behave correctly, and as a programmer would
-expect. 
-\subsection{Applications}
-\subsubsection{Blinker (B)}
-This application was built on top of the app-blinker-chain
-example~\cite{blinker} and modified to remove the delays and LED activations, so
-that it just passes data through three tasks until a maximum count is reached,
-and then the thread terminates. This app was primarily used to demonstrate the
-initial functionality of the system, and then to test mutexes. In both cases a
-combination of \texttt{LOG} statements and final output values were used to
-confirm the correct operation of the system. 
-\subsubsection{Cuckoo-RSA (CRSA)}
-This application is a combination of a cuckoo filter that inserts and queries 64
-values a Bloom filter-like data structure~\cite{Chain}, and RSA encryption of message
-using 128-bit keys~\cite{Chain}. Unfortunately this is the only application
-combination where merging the original applications did not produce the same
-final result as running the applications separately. Several bytes at the end of
-a crucial array for RSA were always overwritten when the applications were
-merged, and we were unable to fix the problem. Unit tests running the RSA
-application in single threaded mode through our scheduler and using our
-multi-threading library produce the correct result, but any combination of RSA
-and cuckoo filtering produces the incorrect result. However, RSA does run to
-completion, so we believe the results from C-RSA are still useful. 
-\subsubsection{Cuckoo-Bitcount (CB)}
-This application is a combination of the cuckoo filter described above and a
-application that uses different algorithms to count the number of bits in an
-array of integers~\cite{bitcount}. CB is a correctly working combination of two
-applications that produces the same output as separate executions of the
-unmodified versions of these applications. It expands the reliability of our
-multi-threading library beyond simple applications like Blinker.
-\subsubsection{Cuckoo-Bitcount-Blinker(CBB)}
-CBB combines the cuckoo filter and bit count applications with two threads of
-the blinker application. This results in four active threads at once, which is
-the maximum number of threads our prototype implementation allows. We confirmed
-that all four threads are fairly serviced by the round robin scheduler and they
-correctly start and end.
-
+\section{Results} - Emily
+No really, there will be things here!
+Applications/measurements\\
+Blinker (x2)\\
+RSA+cukoo\\
+No verification\\
 Memory overhead (read off compilation)\\
@@
 \section{Future Work}
-Future work is focused on saving memory by using channel and code reuse
-optimizations when forking new chain programs from current tasks.  When forking
-a new task, a channel may be duplicated, which linearly increases the amount of
-memory necessary for channels with the number of currently running programs.  It
-is possible to reuse channels between tasks if the number of iterations of that
-task are known at compile time and it can be estimated how many are to be
-running simultaneously. A register allocation graph coloring scheme can then be
-repurposed for reusing channels allocated at compile time.  In either case the
-current channel naming scheme must be changed to support channel duplication,
-ether real or simulated.  Currently channels are allocated and named at compile
-time by combining the names of the input and output channel tasks.  However, if
-tasks can be duplicated this naming scheme must be altered to support multiple
-channels of the same name and potentially multiple channels of the same name.
-This can be done with a vector clock timestamp or a simple counter stored in
-nonvolatile memory to create unique channel names.
+Future work is focused on saving memory by using channel and code reuse optimizations when forking new chain programs from current tasks.
+When forking a new task, a channel may be duplicated, which linearly increases the amount of memory necessary for channels with the number of currently running programs.
+It is possible to reuse channels between tasks if the number of iterations of that task are known at compile time and it can be estimated how
+many are to be running simultaneously. A register allocation graph coloring scheme can then be repurposed for reusing channels allocated at compile time.
+In either case the current channel naming scheme must be changed to support channel duplication, ether real or simulated.
+Currently channels are allocated and named at compile time by combining the names of the input and output channel tasks.
+However, if tasks can be duplicated this naming scheme must be altered to support multiple channels of the same name and potentially multiple channels of the same name.
+This can be done with a vector clock timestamp or a simple counter stored in nonvolatile memory to create unique channel names.
 
-Additionally, it is possible to save memory by sharing arrays between tasks
-without the necessity of copying the array in its entirety in each channel.  To
-do this the concept of intermittent tasks mutexes for shared resources and data
-structures would be introduced.  The structures are protected by a lock that
-contains the thread ID of the task that currently holds the lock.  In order to
-allow all tasks access to a data structure without breaking all chain semantics,
-a modified version of a module channel would be used in order for multiple tasks
-to potentially modify one channel.  However, even with that change it is
-challenging to define sharing between threads in an intermittent context and to
-maintain chain semantics while implementing shared data structures while still
-maintaining correctness.  Because channel sizes are constrained at compile time
-currently, it is also necessary to know at compile time all data structures,
-shared or private, that are going to be used during execution of a chain program
-or programs.  It is a challenge for future work in this field to constrain
-memory for channels at compile time and for the compiler to know with certainty
-how much memory is necessary to allocate. Depending on the workload, the amount
-of forked threads would need to be constrained such that threads don't overflow
-nonvolatile memory at runtime.
+Additionally, it is possible to save memory by sharing arrays between tasks without the necessity of copying the array in its entirety in each channel.
+To do this the concept of intermittent tasks mutexes for shared resources and data structures would be introduced.
+The structures are protected by a lock that contains the thread ID of the task that currently holds the lock.
+In order to allow all tasks access to a data structure without breaking all chain semantics, a modified version of a module channel would be used in order
+for multiple tasks to potentially modify one channel.
+However, even with that change it is challenging to define sharing between threads in an intermittent context and to maintain chain semantics while implementing
+shared data structures while still maintaining correctness.
+Because channel sizes are constrained at compile time currently, it is also necessary to know at compile time all data structures, shared or private,
+that are going to be used during execution of a chain program or programs.
+It is a challenge for future work in this field to constrain memory for channels at compile time and for the compiler to know with certainty how much memory
+is necessary to allocate. Depending on the workload, the amount of forked threads would need to be constrained such that threads don't overflow nonvolatile memory at runtime.
 
-Future work also includes optimizations to reduce overhead, such as converting
-the free indices array to a free indices bitmap in order to reduce the amount of
-baseline memory that the scheduler task uses.  Re-assignation and re-use of
-channels at runtime would increase overhead, but it is thought that the memory
-saved by this procedure is worth the trade-off of increased overhead.  If
-re-assignation can be determined at compile-time, the overhead during runtime
-will still remain low.
+Future work also includes optimizations to reduce overhead, such as converting the free indices array to a free indices bitmap in order to reduce the amount of
+baseline memory that the scheduler task uses.
+Re-assignation and re-use of channels at runtime would increase overhead, but it is thought that the memory saved by this procedure is worth the trade-off of increased overhead.
+If re-assignation can be determined at compile-time, the overhead during runtime will still remain low.
 

### 2017-05-11 paper.tex
conclusion

@@
 \section{Conclusion}
-Multithreading in intermittence {\em is} a good idea!
+The increased prevalence of IoT devices necessitates an increased ability to 
+run larger, general purpose applications on devices with smaller memory and
+energy footprints. With the increased research on intermittently powered devices
+for use in the embedded sphere, it is a logical next step to explore the use 
+of general purpose threaded code on an intermittent device. We have shown
+through the design and implementation of TAPIR on top of the Chain intermittent
+programming model that adding functionality for multiprogramming with
+multithreaded applications can have a low enough memory and power overhead to be
+considered useful on such a device. Additionally, we have implemented a safe
+locking structure for use in nonvolatile memory for such applications that 
+maintains RMW restrictions in the intermittent execution model and allows
+for larger shared data structures without the overhead of memory duplication
+that may overcrowd a small device. We have shown with this design that threaded
+applications can be programmed in a familiar way to software designers and that
+it is a viable next step in intermittent device research.
 

### 2017-05-11 paper.tex
removed capybara files, added explicit TODOs

--- 2017-05-11 paper.tex
@@
 
-SIMD execution on an intermittent system relies on nonvolatile memory instead
-of independent registers for each thread. In Chain's case the implementation
-would rely on channels for each thread running in the system. In the MIMD
-execution model channels could potentially be re-assigned to already allocated
-channels because the threads may end at interleaving times. However, in the
-SIMD model it is expected that each identical thread is at a comperable point
-in execution to the other running threads, and so each thread that may be
-run necessitates a new channel allocation. This could lead to an unavoidable
-hard limit on the number of threads that can run in an applicatioin which in
-turn limits potential applications that are useful on this intermittent device.
+SIMD w/ channels (you don't get your own registers...etc -
+    just channels, since intermittent)\\
 
@@
 
-Mutexes (Simple, meant for pseudo channels) - Emily\\
-Interrupts (mucked with IE flags) - Neil\\
+\subsection{Interrupts}
+% TODO Interrupts (mucked with IE flags) - Neil\\
 % Interrupt prologue
 % Interrupt setup
-%\section{Implementation}
-%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
-%Mutexes (Simple, meant for pseudo channels) - Emily\\
-%Interrupts (mucked with IE flags) - Neil\\
-%% Interrupt prologue
-%% Interrupt setup
+
+% TODO Mutexes (Simple, meant for pseudo channels) - Emily\\
 

### 2017-05-11 paper.tex
Added section for interrupt implementation

--- 2017-05-11 paper.tex
@@
 \section{System Implementation } % TODO change title
-Mutexes - Emily\\
-
 TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
@@
 breaks under \chain's intermittence. Task idempotence in \chain only works
-when execution returns to the top of a
-task upon power failure. Program state (registers and stack memory) are stored
+when execution returns to the top of a task upon power failure. Program state
+(registers and stack memory) are stored
 in volatile memory, if an interrupt handler returned to the middle of a task,
@@
 
-    // Index of current running thread in threads[]
+    // Index of current running thread
     // Stored in self channel
@@
 
-Mutexes (Simple, meant for pseudo channels) - Emily\\
-Interrupts (mucked with IE flags) - Neil\\
-% Interrupt prologue
-% Interrupt setup
-%\section{Implementation}
-%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
-%Mutexes (Simple, meant for pseudo channels) - Emily\\
-%Interrupts (mucked with IE flags) - Neil\\
-%% Interrupt prologue
-%% Interrupt setup
+\subsection{Interrupts}
+The primary issue with interrupts is the interrupt enable flags. The MSP430,
+like many other embedded arhitetures has a master interrupt enable,
+as well as specific, device-level interrupt enables. Both are the user's
+responsibility, and need to be specified in the user-defined
+\texttt{INIT\_FUNC}, as both will be cleared on reset. When an interrupt fires,
+however, interrupts will be disabled by the CPU (again, on most architectures)
+and if the device loses power immediately after an interrupt occurs, there will
+be no record that the interrupt occured. This case is unaviodable on
+intermittent power without explicit support from the power system - we reason
+that if the user cared about every interrupt that fired and if missing one
+interrupt would be catestrophic to the device's operation, then, arguably, that
+device should not be intermittently powered.
+
+Beyond this, there remains the issue of how to return to the interrupt task
+with the interrupt enable set to the correct value if a power failure occurs.
+Since interrupts are disabled on boot (on the MSP430, for instance), the user
+needs to add code to their \texttt{INIT\_FUNC} to enable interrupts. However,
+since interrupt handlers are assumed to be non-reentrant on our system, if
+we are inside an interrupt handler, we need to keep interrupts disabled. We
+also need to write the address of the task that we should return to upon boot -
+the address of the \texttt{INTERRUPT\_TASK} on our system. If these are two
+separate variables, there is a correctness issue with respect to power
+failures. If we write the location of the \texttt{INTERRUPT\_TASK}, but die
+before we can write the fact that we are in an interrupt handler, we will
+return to te interrupt handler with interrupts enabled (since they will be
+enabled by the \texttt{INIT\_TASK}. If we reverse the order and write the fact
+that we are in an interrupt handler, then die before we can write the address
+of the \texttt{INTERRUPT\_TASK}, we will resume execution at a non-interrupt
+task with interrupts disabled, since the \texttt{INIT\_TASK} will not re-enable
+them. Both of these cases could cause execution to fail; the former fails if
+the interrupt handler is not written to be reentrant, the latter fails since
+interrupts will never be processed. Thus, these writes need to be atomic.
+
+We could depend on the programmer to designate all tasks that would be run as
+part of the interrupt handler, then use whether the current task address was
+equivalent to one of those as the flag that designates that we are in an
+interrupt handler, but this forces tasks to either be used as a part of the
+interrupt handler or not, which is bad for code reuse. This also prevents the
+user from being able to spawn off non-interrupt tasks as part of an interrupt
+handler.
+
+Instead, we exploit the fact that, generally, architectures are byte-aligned
+with a fixed instruction size of 16 bits or more - this means, at the very
+least, the lowest bit of any code address is zero. As this is true of ARM,
+RISC-V, MSP430, and most RISC architecutres, we feel that this is a reasonable
+assumption. We use this fact to store the flag that is set if execution is
+inside an interrupt handler in the lowest bit of the address of the task that
+we are currently executing (the task that execution will return to upon
+reboot). Thus, we can use a single write to encode all the needed information,
+giving atomicity. For simplicity for the programmer, upon an interrupt firing,
+an \texttt{interrupt\_prologue()} is run that does this write. The address stored
+in the current context is still the address of the user-defined
+\texttt{INTERRUPT\_TASK}. In our model, this write is the guarentuee that the
+system has recieved the interrupt - if power fails between when the interrupt
+fires and this write occurs, the interrupt will be ignored. This could be fixed
+by having a special capacitor in the power system to accomidate the write, but
+assumptions about the power system, arguably, should not be a part of \chain.
+
+To avoid the pathological stack corruption case described in the motivation for
+interrupts, we keep a flag in non-volatile memory that is set if a reboot has
+occured while in an interrupt handler and is unset otherwise. This flag is
+unset in the afformentioned \texttt{interrupt\_prologue()} and set in
+\texttt{main()}. If and only if the flag is unset when
+\texttt{return\_from\_interrupt()} is called, the two values pushed on the
+stack from an interrupt firing are popped off. Otherwise the stack remains
+unchanged, since the values will have been effectively cleared by a power
+failure.
+
+As of this writing, interrupts are still being debugged, but the above
+reasoning seems, in theory, correct.
+% TODO Mutexes (Simple, meant for pseudo channels) - Emily\\
 

### 2017-05-12 paper.tex
Cleaned up and marked TODOs
abstract

--- 2017-05-11 paper.tex
@@
 \begin{abstract}
-Killer abstract.
+The growth of research in IoT related wireless technologies and low-power computing
+has led to the new area of \textit{energy-harvesting computation} on \textit{intermittently
+powered devices}. Intermittently powered devices are governed by the \textit{intermittent
+execution model} where regular reboots due to an energy charge cycle necessitate division
+of a program into small idempotent \textit{tasks}. However, current applications for
+intermittently powered devices are small and generally contain little functionality.
+
+In this paper we introduce the \textit{Threaded Application Programming on Intermittent Resources} 
+(TAPIR) model to expand the current intermittent programming model \textit{Chain} to 
+include general multiprogramming. TAPIR allows the range of possible programs for 
+intermittent devices to include more general computation functionality while leaving the
+memory and energy overhead low enough that intermittent computation on those workloads 
+still guarantees forward progress. While nonvolatile memory usage will increase per thread
+scheduled, certain optimizations are possible to reduce memory overhead overall. Additionally,
+context switching through our scheduler task is a low-instruction count static overhead that is
+easily accounted for in energy calculations. Through the design of TAPIR we show that general
+purpose threaded applications are feasible on an intermittently powered device.
+
 \end{abstract}
@@
 overhead in scheduling over conventional Chain so that the workload can
-make as much forward progress as possible on one charge cycle.
+make as much forward progress as possible on one charge cycle. While out of
+scope for this project, future work may include a priority scheduler to decrease
+latency for certain threads due to context switching for I/O or other input latency
+sensive workloads.
 

### 2017-05-12 paper.tex

--- 2017-05-11 paper.tex
@@
 
-\section{System Implementation } % TODO change title
+\section{System Implementation}
 TAPIR has a round robin scheduling algorithm. At the end of a task, the compiler
@@
 id, the task will proceed regardless of the fact that \texttt{status = held}.
-%Maybe add a figure here to make the sequence of read/write/failure clearer...
+%TODO Maybe add a figure here to make the sequence of read/write/failure clearer...
 Further, a task must finsih executing before switching to another thread, so

### 2017-05-12 paper.tex
Tweaks to figure

--- 2017-05-12 paper.tex
@@
 \newcommand{\chain}{Chain\xspace}
-
-\usepackage{color}
-
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolour}{rgb}{0.95,0.95,0.95}
-
-\lstdefinestyle{mystyle}{
-    backgroundcolor=\color{backcolour},
-    commentstyle=\color{codegreen},
-    keywordstyle=\color{magenta},
-    numberstyle=\tiny\color{codegray},
-    stringstyle=\color{codepurple},
-    basicstyle=\footnotesize,
-    breakatwhitespace=false,
-    breaklines=true,
-    captionpos=b,
-    keepspaces=true,
-    numbers=none,
-    numbersep=5pt,
-    showspaces=false,
-    showstringspaces=false,
-    showtabs=false,
-    tabsize=4
-    }
-    \lstset{style=mystyle}
-
 \numberofauthors{3}
@@
 \begin{abstract}
-The growth of research in IoT related wireless technologies and low-power computing
-has led to the new area of \textit{energy-harvesting computation} on \textit{intermittently
-powered devices}. Intermittently powered devices are governed by the \textit{intermittent
-execution model} where regular reboots due to an energy charge cycle necessitate division
-of a program into small idempotent \textit{tasks}. However, current applications for
-intermittently powered devices are small and generally contain little functionality.
-
-In this paper we introduce the \textit{Threaded Application Programming on Intermittent Resources} 
-(TAPIR) model to expand the current intermittent programming model \textit{Chain} to 
-include general multiprogramming. TAPIR allows the range of possible programs for 
-intermittent devices to include more general computation functionality while leaving the
-memory and energy overhead low enough that intermittent computation on those workloads 
-still guarantees forward progress. While nonvolatile memory usage will increase per thread
-scheduled, certain optimizations are possible to reduce memory overhead overall. Additionally,
-context switching through our scheduler task is a low-instruction count static overhead that is
-easily accounted for in energy calculations. Through the design of TAPIR we show that general
-purpose threaded applications are feasible on an intermittently powered device.
-
+Killer abstract.
 \end{abstract}
@@
 overhead in scheduling over conventional Chain so that the workload can
-make as much forward progress as possible on one charge cycle. While out of
-scope for this project, future work may include a priority scheduler to decrease
-latency for certain threads due to context switching for I/O or other input latency
-sensive workloads.
+make as much forward progress as possible on one charge cycle.
 
@@
 \begin{minipage}[b]{1.0\textwidth}
-\begin{lstlisting}[language=C]
+\begin{lstlisting}
     /** Enable interrupts if int_setup_complete() has been called
@@
 
-\section{Implementation}
-\subsection{Scheduler}
-\begin{figure*}
-\begin{minipage}[b]{.48\textwidth}
-\begin{lstlisting}[language=C]
-    typedef struct thread_t {
-        unsigned thread_id;
-        context_t context;
-    } thread_t;
-
-    typedef struct thread_state_t {
-        thread_t thread;
-        unsigned active;
-    } thread_state_t;
-
-    // Array of running threads
-    thread_state_t threads[MAX_NUM_THREADS];
-
-    // List of free indicies in threads[]
-    unsigned free_indicies[MAX_NUM_THREADS];
-    unsigned free_indicies_size;
-
-    // Index of current running thread in threads[]
-    // Stored in self channel
-    unsigned current;
-\end{lstlisting}
-\caption{Scheduler Data Structures}\label{schDataStruct}
-\end{minipage}\hfill
-\end{figure*}
-
-The data structures for our scheduler are listed in figure \ref{schDataStruct}.
-Like the rest of chain, memory for data structures is allocated at compile
-time. Thus, the limit on the maximum number of threads allowed in a system is
-fixed arbitrarily - not ideal, but there isn't a better option while operating
-inside chain. The \texttt{threads} array is stored in all-to-one channel, from
-every task to the scheduler task. The \texttt{free\_indicies} array is stored
-in a one-to-all channel, from the scheduler task to every task.
-
-Neither all-to-one nor one-to-all channels fit into standard \chain channel
-semantics, although they resemble multicast-in and multicast-out channels
-described in the original \chain paper. We eschew the notion that channels
-have fixed start and end points and instead treat the channel as a general
-location in non-volatile memory that can be read from and written to by
-arbitrary tasks. However, accesses to these channels are encapsulated by our
-library, so the programmer has no need to think about idempotency violations.
-
-With these multicast channels, the scheduler task can act as a buffer between
-multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
-\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
-\texttt{free\_indicies} through calls to thread create, maintaining
-idempotency. Inside the scheduler task, we read and write \texttt{current},
-so \texttt{current} is double buffered. The only issue is with
-\texttt{thread\_init} since it is called from user-defined tasks and
-initializes \texttt{current} - the same task that called \texttt{thread\_init}
-could call \texttt{thread\_create}, and the \texttt{thread\_create} call
-wouldn't see the update to the variable. Since \texttt{current} is always
-initialized to the same value (0), we opt for manually swapping the buffer on
-\texttt{current} so that other threads can see the update.
-
-We keep an unsigned integer, \texttt{curr\_free\_index}, in volatile memory,
-zeroed on every reboot, that acts as a current pointer in
-\texttt{free\_indicies[]}. Every time the scheduler task runs,
-\texttt{free\_indicies[]} is updated with the state of \texttt{threads[]},
-based on whether or not the active bit is set for a given index. When
-\texttt{thread\_create} is called, if \texttt{curr\_free\_index} is greater
-than \texttt{free\_indicies\_size}, then \texttt{thread\_create} will fail in
-POSIX fashion - the system does not have enough resources to support another
-thread. Otherwise, \texttt{threads[free\_indicies[curr\_free\_index]]} is set
-to the new thread. Upon power failure, \texttt{curr\_free\_index} is reset to
-zero, so as long as the rest of the operations in the task are idempotent, so
-will \texttt{thread\_create} calls - the same locations in \texttt{theads[]}
-will be overwritten.
-
-% TODO picture of array?
-
-\subsection{Interrupts}
-% TODO Interrupts (mucked with IE flags) - Neil\\
-% Interrupt prologue
-% Interrupt setup
-
-% TODO Mutexes (Simple, meant for pseudo channels) - Emily\\
+%\section{Implementation}
+%Scheduler Data structures and idempotence (what we broke and why) -NEIL\\
+%Mutexes (Simple, meant for pseudo channels) - Emily\\
+%Interrupts (mucked with IE flags) - Neil\\
+%% Interrupt prologue
+%% Interrupt setup
 
@@
 expect. 
-
 \subsection{Applications}
@@
 \section{Conclusion}
-The increased prevalence of IoT devices necessitates an increased ability to 
-run larger, general purpose applications on devices with smaller memory and
-energy footprints. With the increased research on intermittently powered devices
-for use in the embedded sphere, it is a logical next step to explore the use 
-of general purpose threaded code on an intermittent device. We have shown
-through the design and implementation of TAPIR on top of the Chain intermittent
-programming model that adding functionality for multiprogramming with
-multithreaded applications can have a low enough memory and power overhead to be
-considered useful on such a device. Additionally, we have implemented a safe
-locking structure for use in nonvolatile memory for such applications that 
-maintains RMW restrictions in the intermittent execution model and allows
-for larger shared data structures without the overhead of memory duplication
-that may overcrowd a small device. We have shown with this design that threaded
-applications can be programmed in a familiar way to software designers and that
-it is a viable next step in intermittent device research.
+Multithreading in intermittence {\em is} a good idea!
 
-\begin{figure}[ht]
+\begin{figure}[hb]
   \centering

### 2017-05-12 paper.tex
Fixed code listings
Added figures for interrupt control flow and the threads array
Spelling fixes

--- 2017-05-12 paper.tex
@@
 purpose threaded applications are feasible on an intermittently powered device.
-
 \end{abstract}
@@
 programs into tasks, which creates automatic checkpointing locations during
-exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks
+execution \cite{Dino}. DINO uses a checkpointing and recovery model that tracks
 both volatile and nonvolatile memory.  The DINO execution model executes all
@@
 Despite the availability of (relatively) inexpensive nonvolatile memory technology such as
-FRAM~\cite{quickrecall}, the the energy and storage overhead of checkpointing multiple
+FRAM~\cite{quickrecall}, the energy and storage overhead of checkpointing multiple
 thread states would limit the maximum per-thread checkpoint size.
@@
 CRFID needs to perform~\cite{dewdrop}. The energy aware memory mapping in ~\cite{Aware}
-discusses the challenge of handling nondeterministic interrupts received by intermittent
+discusses the challenge of handling non-deterministic interrupts received by intermittent
 sensing platforms when a peripheral must be serviced immediately and control flow deviates
@@
 handle {\em multiple} peripherals. Consider a listing in
-~\cite{ufop} reporoduced in Figure ~\ref{fig:copy}, and imagine the level of understanding
+~\cite{ufop} reproduced in Figure ~\ref{fig:copy}, and imagine the level of understanding
 of the hardware behavior a programmer would have to employ to, say, read one sensor if it
@@
 statically defined~\cite{nano}. Contiki uses proto-theads, stackless program threads, to
-offer a model between nano-RK and tinyOS. The Contiki implements "cooperative
+offer a model between Nano-RK and tinyOS. The Contiki implements "cooperative
 multithreading" such that events can trigger a context switch, but multithreading may also
@@
 latency for certain threads due to context switching for I/O or other input latency
-sensive workloads.
+sensitive workloads.
 
@@
 ultimately settled on a multiprogramming model, because this allows for the
-programmer to explictly program unique chain programs with shared or discrete
+programmer to explicitly program unique chain programs with shared or discrete
 data structures on the same system. However, this execution model requires a
@@
 channels because the threads may end at interleaving times. However, in the
-SIMD model it is expected that each identical thread is at a comperable point
+SIMD model it is expected that each identical thread is at a comparable point
 in execution to the other running threads, and so each thread that may be
 run necessitates a new channel allocation. This could lead to an unavoidable
-hard limit on the number of threads that can run in an applicatioin which in
+hard limit on the number of threads that can run in an application which in
 turn limits potential applications that are useful on this intermittent device.
@@
 mutex is set but before the task commits, the task will read out the id of the
-mutex holder before checking if it is free. Upon matchint its id with the holder
+mutex holder before checking if it is free. Upon matching its id with the holder
 id, the task will proceed regardless of the fact that \texttt{status = held}.
 %TODO Maybe add a figure here to make the sequence of read/write/failure clearer...
-Further, a task must finsih executing before switching to another thread, so
-there is no possiblity for an external thread to see a value that was not
+Further, a task must finish executing before switching to another thread, so
+there is no possibility for an external thread to see a value that was not
 committed by another task.
@@
 interface is easier for programmers to use than \texttt{multicast\_channels}, or
-a sepcialized double buffered channel when passing data between threads. It
+a specialized double buffered channel when passing data between threads. It
 maintains familiar semantics for programmers and does not incur the memory
@@
 
+\begin{figure}[ht]
+\begin{minipage}[b]{.9\columnwidth}
+  \includegraphics[width=0.7\columnwidth]{figs/int_ctrl_flow}
+  \caption{Interrupt Control Flow}\label{intCtrlFlow}
+\end{minipage}
+\end{figure}
+
 For the sake of simplicity, we assume that interrupt handlers are not
@@
 \subsection{Scheduler}
-\begin{figure*}
-\begin{minipage}[b]{.48\textwidth}
+
+The data structures for our scheduler are listed in figure \ref{schDataStruct}.
+Like the rest of chain, memory for data structures is allocated at compile
+time. Thus, the limit on the maximum number of threads allowed in a system is
+fixed arbitrarily - not ideal, but there isn't a better option while operating
+inside chain. The \texttt{threads} array is stored in all-to-one channel, from
+every task to the scheduler task. The \texttt{free\_indices} array is stored
+in a one-to-all channel, from the scheduler task to every task.
+
+Neither all-to-one nor one-to-all channels fit into standard \chain channel
+semantics, although they resemble multicast-in and multicast-out channels
+described in the original \chain paper. We eschew the notion that channels
+have fixed start and end points and instead treat the channel as a general
+location in non-volatile memory that can be read from and written to by
+arbitrary tasks. However, accesses to these channels are encapsulated by our
+library, so the programmer has no need to think about idempotency violations.
+
+With these multicast channels, the scheduler task can act as a buffer between
+multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
+\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
+\texttt{free\_indices} through calls to thread create, maintaining
+idempotency. Inside the scheduler task, we read and write \texttt{current},
+so \texttt{current} is double buffered. The only issue is with
+\texttt{thread\_init} since it is called from user-defined tasks and
+initializes \texttt{current} - the same task that called \texttt{thread\_init}
+could call \texttt{thread\_create}, and the \texttt{thread\_create} call
+wouldn't see the update to the variable. Since \texttt{current} is always
+initialized to the same value (0), we opt for manually swapping the buffer on
+\texttt{current} so that other threads can see the update.
+
+\begin{figure}[ht]
+\begin{minipage}[b]{.9\columnwidth}
 \begin{lstlisting}[language=C]
@@
 
-    // List of free indicies in threads[]
-    unsigned free_indicies[MAX_NUM_THREADS];
-    unsigned free_indicies_size;
+    // List of free indices in threads[]
+    unsigned free_indices[MAX_NUM_THREADS];
+    unsigned free_indices_size;
 
@@
 \caption{Scheduler Data Structures}\label{schDataStruct}
-\end{minipage}\hfill
-\end{figure*}
-
-The data structures for our scheduler are listed in figure \ref{schDataStruct}.
-Like the rest of chain, memory for data structures is allocated at compile
-time. Thus, the limit on the maximum number of threads allowed in a system is
-fixed arbitrarily - not ideal, but there isn't a better option while operating
-inside chain. The \texttt{threads} array is stored in all-to-one channel, from
-every task to the scheduler task. The \texttt{free\_indicies} array is stored
-in a one-to-all channel, from the scheduler task to every task.
-
-Neither all-to-one nor one-to-all channels fit into standard \chain channel
-semantics, although they resemble multicast-in and multicast-out channels
-described in the original \chain paper. We eschew the notion that channels
-have fixed start and end points and instead treat the channel as a general
-location in non-volatile memory that can be read from and written to by
-arbitrary tasks. However, accesses to these channels are encapsulated by our
-library, so the programmer has no need to think about idempotency violations.
-
-With these multicast channels, the scheduler task can act as a buffer between
-multiple threads - arbitrary tasks write \texttt{threads[]} with calls to
-\texttt{thread\_create} and \texttt{thread\_end} and tasks read from
-\texttt{free\_indicies} through calls to thread create, maintaining
-idempotency. Inside the scheduler task, we read and write \texttt{current},
-so \texttt{current} is double buffered. The only issue is with
-\texttt{thread\_init} since it is called from user-defined tasks and
-initializes \texttt{current} - the same task that called \texttt{thread\_init}
-could call \texttt{thread\_create}, and the \texttt{thread\_create} call
-wouldn't see the update to the variable. Since \texttt{current} is always
-initialized to the same value (0), we opt for manually swapping the buffer on
-\texttt{current} so that other threads can see the update.
+\end{minipage}
+\end{figure}
 
@@
 zeroed on every reboot, that acts as a current pointer in
-\texttt{free\_indicies[]}. Every time the scheduler task runs,
-\texttt{free\_indicies[]} is updated with the state of \texttt{threads[]},
+\texttt{free\_indices[]}. Every time the scheduler task runs,
+\texttt{free\_indices[]} is updated with the state of \texttt{threads[]},
 based on whether or not the active bit is set for a given index. When
 \texttt{thread\_create} is called, if \texttt{curr\_free\_index} is greater
-than \texttt{free\_indicies\_size}, then \texttt{thread\_create} will fail in
+than \texttt{free\_indices\_size}, then \texttt{thread\_create} will fail in
 POSIX fashion - the system does not have enough resources to support another
-thread. Otherwise, \texttt{threads[free\_indicies[curr\_free\_index]]} is set
+thread. Otherwise, \texttt{threads[free\_indices[curr\_free\_index]]} is set
 to the new thread. Upon power failure, \texttt{curr\_free\_index} is reset to
 zero, so as long as the rest of the operations in the task are idempotent, so
-will \texttt{thread\_create} calls - the same locations in \texttt{theads[]}
+will \texttt{thread\_create} calls - the same locations in \texttt{threads[]}
 will be overwritten.
 
-% TODO picture of array?
+\begin{figure}[ht]
+\begin{minipage}[b]{.9\columnwidth}
+  \includegraphics[width=0.7\columnwidth]{figs/threadsArray}
+  \caption{Threads Array}\label{threadsArray}
+\end{minipage}
+\end{figure}
 
@@
 The primary issue with interrupts is the interrupt enable flags. The MSP430,
-like many other embedded arhitetures has a master interrupt enable,
+like many other embedded architectures has a master interrupt enable,
 as well as specific, device-level interrupt enables. Both are the user's
@@
 and if the device loses power immediately after an interrupt occurs, there will
-be no record that the interrupt occured. This case is unaviodable on
+be no record that the interrupt occurred. This case is unavoidable on
 intermittent power without explicit support from the power system - we reason
 that if the user cared about every interrupt that fired and if missing one
-interrupt would be catestrophic to the device's operation, then, arguably, that
+interrupt would be catastrophic to the device's operation, then, arguably, that
 device should not be intermittently powered.
@@
 before we can write the fact that we are in an interrupt handler, we will
-return to te interrupt handler with interrupts enabled (since they will be
+return to the interrupt handler with interrupts enabled (since they will be
 enabled by the \texttt{INIT\_TASK}. If we reverse the order and write the fact
@@
 least, the lowest bit of any code address is zero. As this is true of ARM,
-RISC-V, MSP430, and most RISC architecutres, we feel that this is a reasonable
+RISC-V, MSP430, and most RISC architectures, we feel that this is a reasonable
 assumption. We use this fact to store the flag that is set if execution is
@@
 in the current context is still the address of the user-defined
-\texttt{INTERRUPT\_TASK}. In our model, this write is the guarentuee that the
-system has recieved the interrupt - if power fails between when the interrupt
+\texttt{INTERRUPT\_TASK}. In our model, this write is the guarantee that the
+system has received the interrupt - if power fails between when the interrupt
 fires and this write occurs, the interrupt will be ignored. This could be fixed
-by having a special capacitor in the power system to accomidate the write, but
+by having a special capacitor in the power system to accommodate the write, but
 assumptions about the power system, arguably, should not be a part of \chain.
@@
 interrupts, we keep a flag in non-volatile memory that is set if a reboot has
-occured while in an interrupt handler and is unset otherwise. This flag is
-unset in the afformentioned \texttt{interrupt\_prologue()} and set in
+occurred while in an interrupt handler and is unset otherwise. This flag is
+unset in the aforementioned \texttt{interrupt\_prologue()} and set in
 \texttt{main()}. If and only if the flag is unset when
@@
 several existing Chain applications that do not share any tasks or channels to
-run concurrently. The programs were run on either an msp430fr5969 launchpad or
+run concurrently. The programs were run on either an MSP430FR5949 launchpad or
 on a Capybara sensing node, a custom intermittent development board with an
-msp430fr5949 MCU. Reboot tests were carried out strictly on the Capybara board
+MSP430FR5949 MCU. Reboot tests were carried out strictly on the Capybara board
 because was simpler to guarantee the isolation of the MCU from the programming
-fet than on the launchpad. We did not seekt to fully verify our implementation
+FET than on the launchpad. We did not seek to fully verify our implementation
 of multi-threading, but we use our experience in application development to
@@
 
-\begin{figure}[ht]
+\begin{figure}[hb]
   \centering

### 2017-05-12 paper.tex
Added authors
Added results

@@
 
-%\author{
-
 % The command \alignauthor (no curly braces needed) should
@@
 
-%\alignauthor Amanda Marano \\
-%        \affaddr{Department of Electrical and Computer Engineering}\\
-%        \affaddr{Carnegie Mellon University}\\
-%       \email{amarano@andrew.cmu.edu}
-%\alignauthor Emily Ruppel \\
-%        \affaddr{Department of Electrical and Computer Engineering}\\
-%        \affaddr{Carnegie Mellon University}\\
-%       \email{eruppel@andrew.cmu.edu}
-%\alignauthor Neil Ryan \\
-%        \affaddr{Department of Electrical and Computer Engineering}\\
-%        \affaddr{Carnegie Mellon University}\\
-%       \email{nryan@andrew.cmu.edu}
-%}
+\author{
+\alignauthor Amanda Marano \\
+        \affaddr{Department of Electrical and Computer Engineering}\\
+        \affaddr{Carnegie Mellon University}\\
+       \email{amarano@andrew.cmu.edu}
+\alignauthor Emily Ruppel \\
+        \affaddr{Department of Electrical and Computer Engineering}\\
+        \affaddr{Carnegie Mellon University}\\
+       \email{eruppel@andrew.cmu.edu}
+\alignauthor Neil Ryan \\
+        \affaddr{Department of Electrical and Computer Engineering}\\
+        \affaddr{Carnegie Mellon University}\\
+       \email{nryan@andrew.cmu.edu}
+}
 
@@
 that it just passes data through three tasks until a maximum count is reached,
-and then the thread terminates. This app was primarily used to demonstrate the
+and then the thread terminates. This app was used to demonstrate the
 initial functionality of the system, and then to test mutexes. In both cases a
 combination of \texttt{LOG} statements and final output values were used to
-confirm the correct operation of the system.
+confirm the correct operation of the system, however we do not report memory
+usage or reboots during execution for this application. 
 \subsubsection{Cuckoo-RSA (CRSA)}
@@
 and cuckoo filtering produces the incorrect result. However, RSA does run to
-completion, so we believe the results from C-RSA are still useful.
+completion, so we believe at least the memory overhead results from C-RSA are still useful.
 \subsubsection{Cuckoo-Bitcount (CB)}
@@
 
-Memory overhead (read off compilation)\\
-Performance (\# reboots)\\
+\subsection{Results}
+CRSA, CB, and CBB were ported from multithreaded applications to serial
+applications. Rather than using the scheduler to switch between tasks, the an
+application transitions to the first task of the successive application upon
+completion. The memory overhead of RSA and Bitcount when they are instrumented
+to run as a single thread with the additional overhead of the scheduler is shown
+to provide a reference for the baseline cost of the scheduler.
+Figure~\ref{fig:mem} shows the increase in code size with the switcht to the
+multithreading library. Approximately 3 lines or code in user space needed to be
+added to each of the multithreaded applications, but the library code is
+significant. The fraction of overhead ranges from 12.6\% for CB to nearly 30\%
+for bitcount.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.99\columnwidth]{figs/memory}
+  \caption{{\bf Increase in code size}}
+  \label{fig:copy}
+\end{figure}
+
+We measure the performance of the application in terms of the number of reboots
+that occur before the application completes. This is an accurate measurement of
+performanc in intermittent systems because the majority of the execution time is
+spent recharging the capacitors. In our system set up, as shown in Figure
+~\ref{fig:setup}, the capacitor charging time is consistent at approximately 3.5
+seconds. We only report reboot data for CBB and CB because of the discrepancy in
+CRSA output between the multithreaded and serial versions.
+Figure~\ref{fig:reboots} shows that the scheduler has a significant cost to the
+system. The multithreaded executions incur significantly more reboots than the
+serial versions. This is to be expected due to the additional instructions of
+scheduler code and additional writes to non-volatile memory required to access
+the scheduler. 
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.99\columnwidth]{figs/setup}
+  \caption{{\bf Test setup using a Capybara board, and RF power source}}
+  \label{fig:setup}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=0.99\columnwidth]{figs/reboots}
+  \caption{{\bf Number of reboots per execution}}
+  \label{fig:reboots}
+\end{figure}
 
